---
title: 'Code 1: Inidividual Radiocarbon Calibration and Correction'
author: "Bryan Rodriguez"
date: "2026-03-25"
output: html_document
---

***Disclaimer: This code was designed by Bryan Rodriguez-Colon with the help of chatGPT-5.***


```{r}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
#setwd('/Users/bryanrodriguez/Library/Mobile Documents/com~apple~CloudDocs/PhD-PROJECT/R-codes/PRGeobio_XR_C14_StableIsotopes_Codes/RadioCarbon_Codes')
```


```{r}
library(IntCal) #For radiocarbon calibration.
library(rbacon) #For age-modeling.
library(clam)
library(coffee)
library(ggplot2)
library(ggsci)
library(RColorBrewer)
library(plotly)
library(dplyr)
library(tibble)
library(purrr)
library(ggpubr)
library(lubridate)
library(patchwork)
library(baseline)
library(EEM)
library(data.table)
library(tidyverse)
library(conflicted)
library(stringr)
library(svglite)
library(ragg)
```



```{r}

# ---- BP -> calendar year helpers ----
bp_to_year_ce <- function(bp) 1950 - bp
fmt_ce_bce    <- function(year_ce)
  ifelse(year_ce >= 1, sprintf("%d CE", round(year_ce)),
                    sprintf("%d BCE", abs(round(year_ce - 1))))

# ---- Calibrate without drawing (for summaries) ----
calibrate_noplot <- function(...) {
  grDevices::pdf(NULL); on.exit(try(grDevices::dev.off(), silent = TRUE), add = TRUE)
  IntCal::calibrate(...)
}

# ---- CalOne: summarized individual calibration (same for C2 & C3) ----
CalOne <- function(age, err, cc, reservoir = 0, res.err = 0, prob = 0.95){
  res  <- calibrate_noplot(age = age, error = err, cc = cc,
                           reservoir = reservoir, resError = res.err, prob = prob)

  dens <- as.data.frame(res[[1]]); names(dens) <- c("calBP","post")
  dens$post <- pmax(dens$post, 0)
  dens <- aggregate(post ~ calBP, data = dens, FUN = sum)
  w <- dens$post / sum(dens$post)

  mean_cal   <- sum(dens$calBP * w)
  cw         <- cumsum(w)
  median_cal <- dens$calBP[which(cw >= 0.5)[1]]

  hpd <- as.data.frame(res[[2]]); names(hpd)[1:2] <- c("from","to")
  hpd_low  <- min(pmin(hpd$from, hpd$to))
  hpd_high <- max(pmax(hpd$from, hpd$to))

  list(median = as.numeric(median_cal),
       mean   = as.numeric(mean_cal),
       hpd95_low  = as.numeric(hpd_low),
       hpd95_high = as.numeric(hpd_high))
}

# ---- Choose curve by material (your ΔR=137±20 for marine) ----
summarize_row <- function(type, age, err, dR = 137, dR_sd = 20){
  if(type %in% c("Plant/Wood","Peat")){
    CalOne(age, err, cc = 1)
  } else if(type == "Sediment Organic carbon"){
    CalOne(age, err, cc = 2, reservoir = dR, res.err = dR_sd)
  } else {
    stop(paste("Unknown Type:", type))
  }
}

# ---- Save the IntCal-style plot for one calibration result ----
# This uses IntCal's own plot method, so the figure matches your first screenshot.
save_intcal_plot <- function(res, file_base, width_in = 5, height_in = 3, dpi = 600){
  # SVG (vector)
  svglite::svglite(paste0(file_base, ".svg"), width = width_in, height = height_in)
  plot(res)                 # IntCal's plot method
  dev.off()

  # TIFF (publication raster)
  ragg::agg_tiff(filename = paste0(file_base, ".tiff"),
                 width = width_in, height = height_in, units = "in",
                 res = dpi, compression = "lzw")
  plot(res)
  dev.off()
}

# ---- Build an IntCal calibration result for one row (for plotting) ----
calibrate_for_row <- function(type, age, err, dR = 137, dR_sd = 20, prob = 0.95){
  if(type %in% c("Plant/Wood","Peat")){
    IntCal::calibrate(age = age, error = err, cc = 1, prob = prob)
  } else if(type == "Sediment Organic carbon"){
    IntCal::calibrate(age = age, error = err, cc = 2,
                      reservoir = dR, resError = dR_sd, prob = prob)
  } else stop(paste("Unknown Type:", type))
}

```



```{r}

# force IntCal's S3 method (works even if generic dispatch fails)
plot_intcal <- function(x, ...) {
  getS3method("plot", "calibration", asNamespace("IntCal"))(x, ...)
}

save_intcal_plot <- function(res, file_base, width_in = 5, height_in = 3, dpi = 600){
  # SVG (vector)
  svglite::svglite(paste0(file_base, ".svg"), width = width_in, height = height_in)
  plot_intcal(res)
  dev.off()

  # TIFF (publication raster)
  ragg::agg_tiff(filename = paste0(file_base, ".tiff"),
                 width = width_in, height = height_in, units = "in",
                 res = dpi, compression = "lzw")
  plot_intcal(res)
  dev.off()
}

calibrate_for_row <- function(type, age, err, dR = 137, dR_sd = 20, prob = 0.95){
  if (type %in% c("Plant/Wood","Peat")) {
    IntCal::calibrate(age = age, error = err, cc = 1, prob = prob)
  } else if (type == "Sediment Organic carbon") {
    IntCal::calibrate(age = age, error = err, cc = 2,
                      reservoir = dR, resError = dR_sd, prob = prob)
  } else stop(paste("Unknown Type:", type))
}

# example loop for C2
outdir <- "figs_calibrations_C2"
if (!dir.exists(outdir)) dir.create(outdir, recursive = TRUE)

for (i in seq_len(nrow(C2_Data))) {
  r <- C2_Data[i, ]
  if (!r$Type %in% c("Plant/Wood","Peat","Sediment Organic carbon")) next
  res  <- calibrate_for_row(r$Type, r$Age, r$Age_Err, dR = 137, dR_sd = 20)
  base <- file.path(outdir, sprintf("%s_%03dcm_%s",
                                    r$Core, as.integer(r$Depth), gsub("[ /]", "-", r$Type)))
  save_intcal_plot(res, base, width_in = 5, height_in = 3, dpi = 600)
}


```








# Functions

## 'CalOne()'

Calibrate a single radiocarbon date and summarize the posterior

description:
Wrapper around `IntCal::calibrate()` that returns **posterior median, posterior mean, and the overall 95% HPD range** (union of HPD segments), all in **calendar years BP (cal BP)**. Works for IntCal20 (terrestrial) and Marine20 (with optional regional reservoir correction ΔR ± σ).

@param age Numeric. Conventional radiocarbon age (years BP), e.g. 2030.
@param err Numeric. 1σ measurement uncertainty of the radiocarbon age, e.g. 15.
@param cc Integer. Calibration curve:
      - `1` = IntCal20 (terrestrial),
      - `2` = Marine20,
      - `3` = SHCal20 (southern hemisphere) [supported by `IntCal::calibrate()`].
      
@param reservoir Numeric. Regional reservoir correction ΔR (years) for marine (or 0 for none). Ignored if `cc != 2`.
@param res.err Numeric. 1σ uncertainty of ΔR (years). Ignored if `cc != 2`.
@param prob Numeric in (0,1). Credible mass for HPD; default `0.95` (2σ).

@return A named list with:
\describe{
\item{median}{Posterior median calendar age (cal BP).}
\item{mean}{Posterior mean calendar age (cal BP).}
\item{hpd95_low}{Lower bound of the **overall** HPD union (min over segments).}
\item{hpd95_high}{Upper bound of the **overall** HPD union (max over segments).}
}

@details
- The function computes **mean/median from the full calibrated pdf** (not from a single HPD slice), which avoids bias near calibration-curve wiggles or multi-modal posteriors.

- The HPD returned is the **union envelope**: the minimum of all segment starts and the maximum of all segment ends. If you need each HPD segment separately, use `IntCal::calibrate(...)[[2]]` directly.

- Output is always in **cal BP**.

@section Dependencies:
Requires `IntCal` package (uses `IntCal::calibrate()`).

@examples
Terrestrial (IntCal20): 1000 ± 15 14C BP
CalOne(age = 1000, err = 15, cc = 1)

Marine (Marine20) with ΔR = 137 ± 20
CalOne(age = 2030, err = 15, cc = 2, reservoir = 137, res.err = 20)



```{r CalOne function, include= T, echo= T}

# To stop images from being created after calling calibrate()
calibrate_noplot <- function(...){
  grDevices::pdf(NULL)                     # open a null device (no file created)
  on.exit(try(grDevices::dev.off(), silent = TRUE), add = TRUE)
  IntCal::calibrate(...)
}



CalOne <- function(age, err, cc, reservoir = 0, res.err = 0, prob = 0.95){
  #res  <- IntCal::calibrate(age = age, error = err, cc = cc, reservoir = reservoir, resError = res.err, prob = prob)
res  <- calibrate_noplot(age = age, error = err, cc = cc,
                         reservoir = reservoir, resError = res.err, prob = prob)


  # Density table -> posterior weights
  dens <- as.data.frame(res[[1]])
  names(dens) <- c("calBP","post")
  dens$post <- pmax(dens$post, 0)
  dens <- stats::aggregate(post ~ calBP, data = dens, sum)  # collapse duplicates
  w <- dens$post / sum(dens$post)

  # Mean & median from the full posterior
  mean_cal <- sum(dens$calBP * w)
  cw <- cumsum(w)
  median_cal <- dens$calBP[which(cw >= 0.5)[1]]

  # HPD union (min/max across all segments)
  hpd <- as.data.frame(res[[2]])
  names(hpd)[1:2] <- c("from","to")
  hpd_low  <- min(pmin(hpd$from, hpd$to))
  hpd_high <- max(pmax(hpd$from, hpd$to))

  list(median = as.numeric(median_cal),
       mean   = as.numeric(mean_cal),
       hpd95_low  = as.numeric(hpd_low),
       hpd95_high = as.numeric(hpd_high))
}


```

## Wrapper helper function

```{r Wrapper, include= T, echo= T}
# --- Row-wise wrapper to pick the right curve/ΔR ----------------------------
summarize_row <- function(type, age, err, dR = 137, dR_sd = 20){
  if(type %in% c("Plant/Wood","Peat")){
    CalOne(age, err, cc = 1)
  } else if(type == "Sediment Organic carbon"){
    CalOne(age, err, cc = 2, reservoir = dR, res.err = dR_sd)
  } else {
    stop(paste("Unknown Type:", type))
  }
}
```


## Helper functions for age conversions

```{r helper to convert cal BP → CE/BCE labels}
# --- helper to convert cal BP → CE/BCE labels ---
bp_to_year_ce   <- function(bp) 1950 - bp
fmt_ce_bce      <- function(year_ce) ifelse(year_ce >= 1,
                                            sprintf("%d CE", round(year_ce)),
                                            sprintf("%d BCE", abs(round(year_ce - 1))))

```

## Function to save calibration plots

```{r}

# given an IntCal::calibrate() result, build a ggplot
plot_calibration_pdf <- function(res, title = NULL){
  dens <- as.data.frame(res[[1]]); names(dens) <- c("calBP","post")
  dens$post <- pmax(dens$post, 0)
  # scale to area 1 (optional, for comparable heights)
  dens$post <- dens$post / max(dens$post, na.rm = TRUE)

  # HPD segments (can be multiple rows)
  hpd <- as.data.frame(res[[2]])
  names(hpd)[1:2] <- c("from","to")

  p <- ggplot(dens, aes(x = calBP, y = post)) +
    geom_area(alpha = 0.25) +
    geom_segment(data = hpd,
                 aes(x = from, xend = to, y = 0, yend = 0),
                 linewidth = 2) +
    scale_x_reverse() +
    labs(x = "cal yr BP", y = "Posterior density", title = title) +
    theme_classic()
  p
}

# save both TIFF and SVG with consistent sizes
save_plot_both <- function(p, file_base, width = 4.0, height = 2.0, dpi = 600){
  ggplot2::ggsave(paste0(file_base, ".svg"), plot = p,
                  device = svglite::svglite, width = width, height = height, units = "in")
  ggplot2::ggsave(paste0(file_base, ".tiff"), plot = p,
                  device = ragg::agg_tiff, width = width, height = height,
                  units = "in", dpi = dpi, compression = "lzw")
}


```


# Load Sediment Core data.

```{r load data,  include=T, echo=T}

C14RawData <- read.csv("Raw_Radiocarbon_Data.csv")

#Data by sediment core. 
C2_Data <- subset(C14RawData, Core == "C2")
C3_Data <- subset(C14RawData, Core == "C3")

#Change the Integer classes to numeric classes
IntergerColumns <- c("Depth", "Age", "Age_Err")
C2_Data <- C2_Data %>% mutate_if(names(.) %in% IntergerColumns, as.numeric)
C3_Data <- C3_Data %>% mutate_if(names(.) %in% IntergerColumns, as.numeric)

#Sort the data in numeric order based on the depth column
C2_Data <- C2_Data %>% arrange(C2_Data$Depth)
C3_Data <- C3_Data %>% arrange(C3_Data$Depth)


```


```{r LPC2 Individual Calibrations, include= T, echo= T}

# --- C2 summary with CE/BCE fields (individual calibrations) ---
C2_clean <- C2_Data %>%
  arrange(Depth) %>%
  rowwise() %>%
  mutate(cal = list(summarize_row(Type, Age, Age_Err))) %>%
  ungroup() %>%
  tidyr::unnest_wider(col = cal) %>%
  mutate(
    Curve = dplyr::case_when(
      Type %in% c("Plant/Wood","Peat") ~ "IntCal20",
      Type == "Sediment Organic carbon" ~ "Marine20 (ΔR=137±20)"
    ),
    # rounded manuscript-friendly numbers
    cal_median = round(median),
    cal95_low  = round(hpd95_low),
    cal95_high = round(hpd95_high),
    cal_mean   = round(mean),
    Calibrated_95_range = sprintf("%d–%d", cal95_low, cal95_high),
    Label_median_95HPD  = sprintf("%d [%d–%d]", cal_median, cal95_low, cal95_high),

    # CE/BCE conversions
    year_ce_median = bp_to_year_ce(cal_median),
    year_ce_min95  = bp_to_year_ce(cal95_high),  # (older BP → smaller CE)
    year_ce_max95  = bp_to_year_ce(cal95_low),
    label_median_cal = fmt_ce_bce(year_ce_median)
  ) %>%
  select(Core, Depth, Type, Age, Age_Err, Curve,
         cal_median, cal95_low, cal95_high, cal_mean,
         Calibrated_95_range, Label_median_95HPD,
         year_ce_median, year_ce_min95, year_ce_max95, label_median_cal)

write.csv(C2_clean, "CalibratedDataC2_DeltaR137.csv", row.names = FALSE)

```


## LPC3 core

```{r LPC3 individual calibration, include= T, echo= T}

C3_clean <- C3_Data %>%
  arrange(Depth) %>%
  rowwise() %>%
  mutate(cal = list(summarize_row(Type, Age, Age_Err))) %>%
  ungroup() %>%
  tidyr::unnest_wider(col = cal) %>%
  mutate(
    Curve = dplyr::case_when(
      Type %in% c("Plant/Wood","Peat") ~ "IntCal20",
      Type == "Sediment Organic carbon" ~ "Marine20 (ΔR=137±20)"
    ),
    cal_median = round(median),
    cal95_low  = round(hpd95_low),
    cal95_high = round(hpd95_high),
    cal_mean   = round(mean),
    Calibrated_95_range = sprintf("%d–%d", cal95_low, cal95_high),
    Label_median_95HPD  = sprintf("%d [%d–%d]", cal_median, cal95_low, cal95_high),

    year_ce_median = bp_to_year_ce(cal_median),
    year_ce_min95  = bp_to_year_ce(cal95_high),
    year_ce_max95  = bp_to_year_ce(cal95_low),
    label_median_cal = fmt_ce_bce(year_ce_median)
  ) %>%
  select(Core, Depth, Type, Age, Age_Err, Curve,
         cal_median, cal95_low, cal95_high, cal_mean,
         Calibrated_95_range, Label_median_95HPD,
         year_ce_median, year_ce_min95, year_ce_max95, label_median_cal)

write.csv(C3_clean, "CalibratedDataC3_DeltaR137.csv", row.names = FALSE)


```

## Saving Calibration plots for Dissertation Apendix

```{r}

# where to write
outdir <- "figs_calibrations_C2"
if (!dir.exists(outdir)) dir.create(outdir, recursive = TRUE)

for(i in seq_len(nrow(C2_Data))){
  r <- C2_Data[i, ]
  # skip unknown types
  if(! r$Type %in% c("Plant/Wood","Peat","Sediment Organic carbon")) next

  # build the IntCal calibration object
  res <- calibrate_for_row(r$Type, r$Age, r$Age_Err, dR = 137, dR_sd = 20, prob = 0.95)

  # base name like "C2_055cm_Plant-Wood"
  base <- file.path(
    outdir,
    sprintf("%s_%03dcm_%s", r$Core, as.integer(r$Depth), gsub("[ /]", "-", r$Type))
  )

  # save as SVG + TIFF (green curve + grey pdf, HPD bars)
  save_intcal_plot(res, base, width_in = 5, height_in = 3, dpi = 600)
}

# repeat for C3
outdir <- "figs_calibrations_C3"
if (!dir.exists(outdir)) dir.create(outdir, recursive = TRUE)
for(i in seq_len(nrow(C3_Data))){
  r <- C3_Data[i, ]
  if(! r$Type %in% c("Plant/Wood","Peat","Sediment Organic carbon")) next
  res <- calibrate_for_row(r$Type, r$Age, r$Age_Err, dR = 137, dR_sd = 20, prob = 0.95)
  base <- file.path(outdir, sprintf("%s_%03dcm_%s", r$Core, as.integer(r$Depth), gsub("[ /]", "-", r$Type)))
  save_intcal_plot(res, base, width_in = 5, height_in = 3, dpi = 600)
}

```



