---
title: 'Code 2: Bacon and Sed Rates'
author: "Bryan Rodriguez"
date: "2026-03-25"
output: html_document
editor_options: 
  markdown: 
    wrap: 72
---

***Disclaimer: This code was designed by Bryan Rodriguez-Colon with the
help of chatGPT-5.***

# Code 2: Age–Depth Modeling and Sedimentation Rates (Bacon)

### Overview

This script performs Bayesian age–depth modeling using the `rbacon`
package (Blaauw and Christen, 2011) and extracts sedimentation rates for
sediment cores LPC2 and LPC3. The workflow includes model execution,
extraction of accumulation rates, statistical summarization, conversion
to sedimentation rates, and visualization.

### Workflow

1.  Run Bacon age–depth models using uncalibrated radiocarbon ages
2.  Extract accumulation rate outputs from Bacon `.out` files
3.  Compute median and 95% credible intervals for each depth section
4.  Convert accumulation rates (yr/cm) to sedimentation rates (mm/yr)
5.  Reconstruct depth intervals based on core length and Bacon
    segmentation
6.  Generate and export sedimentation rate plots

### Key Assumptions

-   Bacon performs internal radiocarbon calibration (IntCal20 by
    default)

-   Accumulation rates are reported in yr/cm and converted as:

    $$
    \text{Sedimentation rate (mm/yr)} = \frac{10}{\text{accumulation (yr/cm)}}
    $$

-   Depth intervals are reconstructed assuming uniform section
    thickness, defined by:

    -   LPC2: 115 cm total depth
    -   LPC3: 95 cm total depth

-   Bacon sections are mapped linearly onto measured core depth

### Output

-   Sedimentation rate summaries (median, 95% CI)
-   Depth-resolved sedimentation rate plots (SVG and PNG)

### Reference

Blaauw, M., and Christen, J.A. (2011). Flexible paleoclimate age–depth
models using an autoregressive gamma process. *Quaternary
Geochronology*, 6, 457–474.

```{r}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
#setwd('/Users/bryanrodriguez/Library/Mobile Documents/com~apple~CloudDocs/PhD-PROJECT/R-codes/PRGeobio_XR_C14_StableIsotopes_Codes/RadioCarbon_Codes')
```

```{r Load packages, include=T, message=F}
library(rbacon)   # Bayesian age–depth modeling
library(ggplot2)  # plotting sedimentation rates
library(dplyr)    # data manipulation
library(tidyr)    # data reshaping
```

# Run Bacon Age–Depth Models

Bayesian age–depth models are generated for each core using the `rbacon`
package.\
The model uses prior information on accumulation rates and divides the
core into equal-thickness sections.

Parameters: - `acc.mean`: prior mean accumulation rate (yr/cm) -
`thick`: thickness of each modeled section (cm) - `d.min`, `d.max`:
minimum and maximum depths of the core - `add.bottom`: ensures model
extends to full core depth

```{r Run Bacon}

Bacon(
  core = "LPC2",        # new run name (don’t overwrite)
  acc.mean = 11,
  thick = 5,               # explicitly define section thickness
  add.bottom = TRUE,
 d.min = 0,
 d.max = 115,
  suggest = FALSE,
  accept.suggestions = FALSE
)

Bacon(
  core = "LPC3",        # new run name (don’t overwrite)
  acc.mean = 11,
  thick = 5,               
  add.bottom = TRUE,
  d.min = 0,
  d.max = 95,
  suggest = FALSE,
  accept.suggestions = FALSE
)

```

# Sedimentation rates from each core

## LPC2

### Extract Accumulation Rates from Bacon Output

Bacon produces `.out` files containing MCMC iterations of accumulation
rates for each depth section.

Processing steps: 1. Read `.out` file 2. Convert to numeric matrix 3.
Remove memory column (first column) 4. Structure as a data frame for
analysis



```{r Extract_SedRates, include =T}

bacon_raw_LPC2 <- readLines("Bacon_runs/LPC2/LPC2_24.out")

bacon_split_LPC2 <- strsplit(bacon_raw_LPC2, ",")
bacon_mat_LPC2 <- do.call(rbind, lapply(bacon_split_LPC2, as.numeric))

bacon_df_LPC2 <- as.data.frame(bacon_mat_LPC2)

str(bacon_df_LPC2)
```

### Summarize Accumulation Rates

For each depth section: - Median accumulation rate is calculated - 95%
credible intervals (2.5% and 97.5%) are computed

These statistics represent the posterior distribution of sediment
accumulation rates.

### Convert Accumulation to Sedimentation Rates

Accumulation rates (yr/cm) are converted to sedimentation rates (mm/yr):

$$
\text{mm/yr} = \frac{10}{\text{yr/cm}}
$$

This allows direct interpretation of sedimentation intensity through
time.

```{r Summarize, include=T}
acc_df_LPC2 <- bacon_df_LPC2[, -1]  # remove memory column


acc_summary_LPC2 <- acc_df_LPC2 %>%
  summarise(across(everything(), list(
    med = ~median(.x),
    low = ~quantile(.x, 0.025),
    high = ~quantile(.x, 0.975)
  )))


acc_summary_long_LPC2 <- acc_summary_LPC2 %>%
  pivot_longer(everything(),
               names_to = c("section", ".value"),
               names_sep = "_")

acc_summary_long_LPC2 <- acc_summary_long_LPC2 %>%
  mutate(rate_mm_yr = 10 / med)

```


### Reconstruct Depth Intervals

Depth intervals are reconstructed from Bacon sections assuming uniform
thickness.

-   LPC2 total depth: 115 cm

Section thickness is calculated as:

$$
\text{section thickness} = \frac{\text{core depth}}{\text{number of sections} - 2}
$$

The subtraction accounts for edge sections generated by Bacon that are
excluded.

Depth metrics calculated: - Top depth - Bottom depth - Midpoint depth

```{r Depth Intervals, include=T}
section_thickness <- 115 / (26 - 2) #Depth of the core

acc_clean_LPC2 <- acc_summary_long_LPC2 %>%
  mutate(
    section_num = as.numeric(gsub("V", "", section)) - 1
  ) %>%
  filter(!section %in% c("V26", "V27")) %>%
  mutate(
    top_depth_cm = (section_num - 1) * section_thickness,
    bottom_depth_cm = section_num * section_thickness,
    mid_depth_cm = (top_depth_cm + bottom_depth_cm) / 2
  )

acc_clean_LPC2 <- acc_clean_LPC2 %>%
  mutate(
    low_trim = ifelse(low < quantile(low, 0.05), NA, low)
  )
```


## C3

Same process as before, only that:

Depth intervals are reconstructed from Bacon sections assuming uniform
thickness.

-   LPC3 total depth: 95 cm

```{r}

bacon_raw_LPC3 <- readLines("Bacon_runs/LPC3/LPC3_20.out")

bacon_split_LPC3 <- strsplit(bacon_raw_LPC3, ",")
bacon_mat_LPC3 <- do.call(rbind, lapply(bacon_split_LPC3, as.numeric))

bacon_df_LPC3 <- as.data.frame(bacon_mat_LPC3)

str(bacon_df_LPC3)
```

```{r}
acc_df_LPC3 <- bacon_df_LPC3[, -1]  # remove memory column


acc_summary_LPC3 <- acc_df_LPC3 %>%
  summarise(across(everything(), list(
    med = ~median(.x),
    low = ~quantile(.x, 0.025),
    high = ~quantile(.x, 0.975)
  )))


acc_summary_long_LPC3 <- acc_summary_LPC3 %>%
  pivot_longer(everything(),
               names_to = c("section", ".value"),
               names_sep = "_")

acc_summary_long_LPC3 <- acc_summary_long_LPC3 %>%
  mutate(rate_mm_yr = 10 / med)

```

```{r}

n <- length(unique(acc_summary_long_LPC3$section))

section_thickness <- 95 / (n - 2)

acc_clean_LPC3 <- acc_summary_long_LPC3 %>%
  mutate(
    section_num = as.numeric(gsub("V", "", section)) - 1
  ) %>%
  filter(!section %in% c("V22", "V23")) %>%
  mutate(
    top_depth_cm = (section_num - 1) * section_thickness,
    bottom_depth_cm = section_num * section_thickness,
    mid_depth_cm = (top_depth_cm + bottom_depth_cm) / 2
  )

acc_clean_LPC3 <- acc_clean_LPC3 %>%
  mutate(
    low_trim = ifelse(low < quantile(low, 0.05), NA, low)
  )
```


### Plot Sedimentation Rates

Sedimentation rates are visualized as depth-resolved rectangles: -
X-axis: sedimentation rate (mm/yr) - Y-axis: depth (cm, reversed)

Each rectangle represents a modeled depth interval with its
corresponding rate.

```{r plotLPC2SedRates}
plotC2 <- ggplot(acc_clean_LPC2,
       aes(ymin = top_depth_cm, ymax = bottom_depth_cm,
           xmin = 0, xmax = rate_mm_yr)) +
    geom_rect(fill = "grey60") +
    scale_y_reverse() +
    labs(
        x = "Sedimentation rate (mm/yr)",
        y = "Depth (cm)"
    ) +
    theme_classic() +
  ggtitle("LPC2 Sed Rates")
plotC2
```

```{r plotLPC3SedRates, include=T}
plotC3 <- ggplot(acc_clean_LPC3,
       aes(ymin = top_depth_cm, ymax = bottom_depth_cm,
           xmin = 0, xmax = rate_mm_yr)) +
    geom_rect(fill = "grey60") +
    scale_y_reverse() +
    labs(
        x = "Sedimentation rate (mm/yr)",
        y = "Depth (cm)"
    ) +
    theme_classic() +
  ggtitle("LPC3 Sed rates")
plotC3
```

# Save Sedimentation rate plot

Plots are exported in: - SVG format (vector, publication-quality) - PNG
format (raster)

Dimensions are standardized for consistency across figures.

```{r save Plots, include=T}
ggsave(
  filename = "plots/C2/sed_rate_C2.svg",
  plot = plotC2,
  width = 4,
  height = 10,
  units = "in",
  device = "svg"
)

ggsave(
  filename = "plots/C2/sed_rate_C2.png",
  plot = plotC2,
  width = 4,
  height = 10,
  units = "in",
  device = "png"
)

ggsave(
  filename = "plots/C3/sed_rate_C3.svg",
  plot = plotC3,
  width = 4,
  height = 10,
  units = "in",
  device = "svg"
)

ggsave(
  filename = "plots/C3/sed_rate_C3.png",
  plot = plotC3,
  width = 4,
  height = 10,
  units = "in",
  device = "png"
)
```
