---
title: 'Code2: XRD PCA Analysis'
author: "Bryan Rodriguez"
date: "2026-03-17"
output: html_document
---

***Disclaimer: This code was designed by Bryan Rodriguez-Colon with the help of chatGPT-5.***

# Code2_PCA_XRFcores

## Overview

This script performs Principal Component Analysis (PCA) on clr-transformed XRF elemental datasets from sediment cores C2 and C3. The goal is to identify dominant geochemical gradients, evaluate compositional variability, and interpret these patterns within a stratigraphic (facies-based) framework.

This script operates on preprocessed data generated in **Code1_clr_XRFcores**, ensuring a modular workflow where data transformation and statistical analysis are separated.

## Objectives

1. Load clr-transformed datasets  
2. Perform PCA on compositional geochemical data  
3. Visualize sample distributions and elemental controls  
4. Quantify variance structure  
5. Export figures and processed data for interpretation and external workflows  

## Workflow structure

### 1. Data loading

clr-transformed datasets (`clr_data_C2`, `clr_data_C3`) are imported from a saved `.RData` file. These datasets already:

- Contain filtered elements  
- Are free of zeros and missing values  
- Include depth information  

### 2. Data preparation

Prior to PCA:

- Non-compositional variables (e.g., `Depth`) are removed  
- Elements interpreted as analytical or post-processing artifacts (e.g., `K`, `Cl`) are excluded  

This ensures that PCA is performed only on geochemically meaningful clr-transformed variables.

### 3. PCA computation

PCA is performed using the `prcomp()` function with:

- **Centering** (mean subtraction)  
- **Scaling** (unit variance standardization)  

This approach ensures that all variables contribute equally, independent of their variance or magnitude.

### 4. PCA outputs

For each core:

- **Scores (PC1, PC2)**: represent sample positions in reduced-dimensional space  
- **Loadings**: represent the contribution of each element to the principal components  

Depth information is reattached to scores for stratigraphic interpretation.

### 5. Stratigraphic (facies) classification

Samples are grouped into facies based on depth intervals to contextualize PCA results.

#### Core C2 facies

- ZC2A: 0–30 cm  
- ZC2B: 30–48 cm  
- ZC2C: 48–60 cm  
- ZC2D: 60–107 cm  
- ZC2E: 107–116 cm  

#### Core C3 facies

- ZC3A: 0–15 cm  
- ZC3B: 15–38 cm  
- ZC3C: 38–53.6 cm  
- ZC3D: 53.6–94 cm  

These facies divisions allow interpretation of geochemical variability in a stratigraphic and paleoenvironmental context.

### 6. Visualization

Several types of plots are generated:

#### PCA score plots
- PC1 vs PC2 scatter plots  
- Colored by facies  
- Used to identify clustering and trends  

#### Faceted PCA plots
- Separate panels per facies  
- Highlight intra-facies variability  

#### Interactive plots
- Generated using `plotly`  
- Allow exploration of sample-level metadata (e.g., depth)  

#### Loadings (correlation biplots)
- Display eigenvectors only  
- Show relationships among elements  
- Axes fixed to [-1, 1] for interpretability  

#### Scree plots
- Show percentage of variance explained per component  
- Used to assess dimensionality  

### 7. Variance and correlation analysis

- Explained variance is calculated from PCA eigenvalues  
- A correlation matrix of loadings is generated to quantify relationships among elements  

These outputs support interpretation of geochemical controls and element associations.

### 8. Data and figure export

Outputs are saved for reproducibility and integration with external workflows:

#### Figures
- PCA plots (SVG and TIFF formats)  
- Scree plots  

#### Data
- PCA scores  
- clr-transformed datasets  
- Explained variance  
- Loadings correlation matrices  

These outputs are used for further visualization (e.g., MATLAB-based downcore plots) and publication figures.

### 9. Workspace management

Intermediate variables and large objects are removed after export to:

- Optimize memory usage  
- Avoid object conflicts  
- Maintain a clean analytical environment  

## Notes

- PCA is performed on clr-transformed data to respect compositional data structure (Aitchison, 1986)  
- Element selection and prior filtering strongly influence PCA results  
- Consistent preprocessing across cores ensures comparability  
- Interpretation should integrate stratigraphy, mineralogy, and depositional context  

## Reference

Aitchison, J. (1986). *The Statistical Analysis of Compositional Data*. Chapman & Hall.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
#setwd('/Users/bryanrodriguez/Library/Mobile Documents/com~apple~CloudDocs/PhD-PROJECT/R-codes/PRGeobio_XR_C14_StableIsotopes_Codes/XRF_Codes')

```

Stop data displayed in scientific notation

```{r stop scipen}
options(scipen = 999)
```

```{r load packages, warning=FALSE}

library(ggplot2)      # plotting (all figures)
library(dplyr)        # data manipulation (select, filter, pipes)

library(compositions) # clr conceptual framework
library(zCompositions) # zero-replacement (even if not used directly, justified)

library(MASS)         # LDA (lda function)
library(ggrepel)      # label placement in biplots

library(plotly)       # interactive plots
```


```{r}
load("RData/clr_XRF_cores.RData")
```

# Core C2

## Principal Component Analysis (PCA) – Core C2

Principal Component Analysis (PCA) is applied to the clr-transformed dataset of Core C2 to identify dominant geochemical gradients and evaluate compositional variability across stratigraphic intervals.

### Data preparation

Prior to PCA, the dataset is subset to exclude:

- `Depth` (non-compositional variable)  
- `K` and `Cl`, which are interpreted as potential artifacts related to drying or salt precipitation processes rather than primary depositional signals  

This step ensures that PCA is performed only on geochemically meaningful clr-transformed variables.

### PCA implementation

PCA is performed using the `prcomp()` function with:

- **Centering**: mean subtraction for each variable  
- **Scaling**: standardization to unit variance  

This approach ensures that all elements contribute equally to the analysis regardless of differences in variance or absolute magnitude.

### PCA scores

The first two principal components (PC1 and PC2) are extracted to represent the dominant axes of variance in the dataset.

- These scores correspond to sample positions in reduced-dimensional space  
- They are used for visualization and interpretation of geochemical trends  

Depth information is reattached to the PCA scores to preserve stratigraphic context.

### Facies classification

Samples are grouped into stratigraphic facies based on depth intervals (from base to top):

- ZC2A: 0–30 cm  
- ZC2B: 30–48 cm  
- ZC2C: 48–60 cm  
- ZC2D: 60–107 cm  
- ZC2E: 107–116 cm  

This classification enables interpretation of PCA results in terms of depositional units and vertical (stratigraphic) geochemical transitions.

### Loadings

PCA loadings (eigenvectors) are extracted and scaled by the standard deviation of each principal component. This produces coordinates compatible with score space for biplot visualization.

- Loadings represent the contribution of each element to the principal components  
- High absolute values indicate strong influence on the corresponding axis  

### Notes

- Optional scaling of loadings to match score magnitude is included but not applied by default  
- The interpretation of PCA results depends on the clr-transformed variable set and prior filtering decisions  
- Removing variables (e.g., K, Cl) changes the compositional basis and should be consistently documented  

Temporary subset data is removed after PCA to maintain a clean workspace.

```{r PCA C2}

# --- Subset Data for PCA ---
# Only include the elements you wish to use for the PCA
subset_data_C2 <- clr_data_C2 %>% dplyr::select(-c("Depth", "K", "Cl")) #Taking pout K and Cl as they are probably artifacts from the drying process. 


# --- Perform PCA on clr-transformed data ---
pca_results_C2 <- prcomp(subset_data_C2, center = TRUE, scale. = TRUE)

# --- Extract PCA Scores ---
scores_C2 <- data.frame(pca_results_C2$x[, 1:2])
colnames(scores_C2) <- c("PC1", "PC2")

# --- Add Depth and Facies Information ---
# Here, we assume you have a depth vector (Depth_cm) from your original data.
# Adjust this accordingly if Depth_cm is stored elsewhere.
scores_C2$Depth <- clr_data_C2$Depth  # Replace 'original_data' with your depth data frame

# Create facies based on depth intervals:
# ZC2A: 107-116 cm, ZC2B: 60-107 cm, ZC2C: 48-60 cm, ZC2D: 30-48 cm, ZC2E: 0-30 cm.
scores_C2$Facies <- cut(scores_C2$Depth,
                     breaks = c(0, 30, 48, 58, 107, 116),
                     labels = c("ZC2A", "ZC2B", "ZC2C", "ZC2D", "ZC2E"),
                     include.lowest = TRUE, right = TRUE)

# --- Extract and scale Loadings ---
loadings_C2 <- pca_results_C2$rotation[, 1:2] * 
  matrix(pca_results_C2$sdev[1:2], nrow = nrow(pca_results_C2$rotation), ncol = 2, byrow = TRUE)
loadings_C2 <- data.frame(loadings_C2)
colnames(loadings_C2) <- c("PC1", "PC2")
loadings_C2$variable <- rownames(loadings_C2)

####Scale Loadings to the same coordinate system as the scores (using the maximum absolute score). OPTIONAL###
#scale_factor <- max(abs(scores_C2$PC1), abs(scores_C2$PC2))
#loadings_scaled_C2 <- loadings_C2
#loadings_scaled_C2$PC1 <- loadings_scaled_C2$PC1 * scale_factor
#loadings_scaled_C2$PC2 <- loadings_scaled_C2$PC2 * scale_factor

rm(subset_data_C2)

```

## PCA visualization – Core C2

### PCA biplot (scores only)

The PCA results for Core C2 are visualized using a scatter plot of the first two principal components (PC1 vs PC2), representing the dominant axes of geochemical variability.

- Each point corresponds to a sample  
- Points are colored by stratigraphic facies  
- Transparency (`alpha = 0.7`) is applied to reduce overplotting  

Depth information is included as an interactive tooltip variable, enabling direct inspection of stratigraphic position in the interactive version of the plot.

### Interactive visualization

An interactive version of the PCA plot is generated using `plotly`:

- Allows exploration of individual samples  
- Displays PC coordinates, facies, and depth on hover  
- Facilitates identification of trends, clusters, and outliers  

This is particularly useful for exploratory data analysis and validation of facies groupings.

### Faceted PCA plots

To better visualize intra-facies variability, the PCA scores are also plotted using faceting:

- Each panel represents a single facies  
- Enables comparison of variance structure within and between facies  
- Helps identify whether geochemical variability is facies-dependent or continuous across depth  

### Notes

- These plots display only PCA scores (samples), not loadings  
- Interpretation should consider both clustering patterns and stratigraphic ordering  
- Faceting can highlight subtle differences that may not be apparent in the combined plot  

```{r Plot C2 Individual Factor Map, echo =F}

# --- Create the PCA Biplot (Individual Factor Map) ---
pcaC2_biplot_facies <- ggplot() +
  # Plot sample scores, colored by facies
  geom_point(data = scores_C2, aes(x = PC1, y = PC2, color = Facies, text = Depth), alpha = 0.7, size = 2) +
  theme_minimal() +
  labs(title = "PCA Biplot with Facies for Core C2",
       x = "PC1", y = "PC2", color = "Facies")

# Print the plot
print(pcaC2_biplot_facies)

interactive_plot_PCA_C2 <- ggplotly(pcaC2_biplot_facies, tooltip = c("x", "y", "color", "text"))
interactive_plot_PCA_C2

# This is to make each zone separate in facets
pcaC2_biplot_facies_facets_C2 <- ggplot() +
  # Plot sample scores, colored by facies
  geom_point(data = scores_C2, aes(x = PC1, y = PC2, color = Facies), alpha = 0.7, size = 2) +
  facet_wrap(~ Facies) +
  theme_minimal() +
  labs(title = "PCA Biplot Faceted by Facies for Core C2",
       x = "PC1", y = "PC2", color = "Facies")

pcaC2_biplot_facies_facets_C2
```

## PCA loadings visualization (correlation biplot) – Core C2

To evaluate the contribution of individual elements to the principal components, a correlation biplot is generated using PCA loadings (eigenvectors) only.

### Representation

- Each vector represents an elemental variable  
- Vector direction indicates the relationship with PC1 and PC2  
- Vector length reflects the strength of the contribution to the principal components  

Elements pointing in similar directions are positively correlated, whereas elements pointing in opposite directions are negatively correlated.

### Scaling and axes

- Loadings are plotted within a fixed coordinate system ranging from −1 to 1 on both axes  
- This standardization facilitates interpretation as a correlation structure between variables and principal components  

### Labeling

Element names are displayed using repelled text labels to improve readability and avoid overlap, allowing clear identification of geochemical drivers.

### Interpretation

- Elements with long vectors exert strong control on the variance captured by PC1 and PC2  
- Clusters of vectors indicate groups of elements with similar geochemical behavior  
- Orthogonal relationships suggest independence between variables  

### Notes

- This plot excludes sample scores to focus exclusively on variable relationships  
- It complements the PCA score plots by providing mechanistic insight into the geochemical gradients identified  

```{r Plot C2 only the eigenvectors without the scores}

# Create the correlation biplot with fixed axes [-1, 1]
pca_biplot_cor_C2 <- ggplot() +
  geom_segment(data = loadings_C2, 
               aes(x = 0, y = 0, xend = PC1, yend = PC2), 
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black") +
  geom_text_repel(data = loadings_C2, 
                  aes(x = PC1, y = PC2, label = variable), 
                  color = "red", size = 4) +
  theme_minimal() +
  labs(title = "Correlation Biplot from clr-transformed XRF Data in C2", 
       x = "PC1", y = "PC2") +
  xlim(-1, 1) +
  ylim(-1, 1)

print(pca_biplot_cor_C2)


```


```{r LDA in C2 (OPTIONAL, DO NOT USE)}

# Assuming clr_data_C2 is your clr-transformed data with a column named "Depth" in mm
# Define facies based on depth intervals in mm:
clr_data_C2$Facies <- cut(clr_data_C2$Depth,
                          breaks = c(0, 30, 48, 60, 107, 116),
                          labels = c("ZC2A", "ZC2B", "ZC2C", "ZC2D", "ZC2E"),
                          include.lowest = TRUE, right = TRUE)

# Check the distribution of facies to ensure each group has observations:
print(table(clr_data_C2$Facies))

# Define the predictor elements for LDA
elements_C2_LDA <- elements_keep_C2

# Prepare the data for LDA by subsetting to the chosen elements and the facies column
lda_data <- clr_data_C2[, c(elements_C2_LDA, "Facies")]

# Perform LDA using the chosen elements to discriminate among facies
lda_model <- lda(Facies ~ ., data = lda_data)
lda_pred <- predict(lda_model)

# Combine the LDA scores with facies information
lda_scores <- data.frame(lda_pred$x)
lda_scores$Facies <- lda_data$Facies

# Plot the first two linear discriminants
lda_plot <- ggplot(lda_scores, aes(x = LD1, y = LD2, color = Facies)) +
  geom_point(size = 2) +
  labs(title = "LDA of Core C2 clr-transformed Data",
       x = "LD1",
       y = "LD2",
       color = "Facies") +
  theme_minimal()

print(lda_plot)



```

## Explained variance and correlation structure – Core C2

### Explained variance

The proportion of variance explained by each principal component is calculated from the eigenvalues of the PCA:

- Eigenvalues are obtained from the squared standard deviations (`sdev²`)  
- Values are normalized by the total variance to obtain relative contributions  
- Results are expressed as percentages  

This quantifies how much of the total geochemical variability is captured by each principal component.

### Scree plot

A scree plot is generated to visualize the distribution of explained variance across principal components:

- Each bar represents a principal component  
- The height corresponds to the percentage of variance explained  
- This plot aids in determining how many components are meaningful to retain  

Typically, the first few components capture the majority of the variance and are used for interpretation.

### Loadings correlation matrix

A correlation matrix is computed using the PCA loadings:

- Based on the rotation matrix (eigenvectors)  
- Represents relationships among elemental variables in PCA space  

This matrix provides complementary information to the biplot by:

- Quantifying the degree of association between variables  
- Identifying groups of elements with similar geochemical behavior  
- Supporting interpretation of element clustering observed in the loadings plot  

### Notes

- The explained variance structure is essential for assessing dimensionality reduction efficiency  
- The correlation matrix reflects relationships in transformed (clr) space, not raw counts  
- Naming consistency between variance and correlation outputs is maintained for clarity in downstream export and interpretation  

```{r Explained variance and Correlation Matrix for C2}
# Explained Variance
explained_variance_C2 <- pca_results_C2$sdev^2 / sum(pca_results_C2$sdev^2)
explained_variance_perc_C2 <- explained_variance_C2 * 100

# Print Explained Variance
print(explained_variance_perc_C2)

# Creating the barplot for Explained Variance
pc_var_C2 <- data.frame(PC = seq(along = explained_variance_perc_C2), Variance = explained_variance_perc_C2)
barplot_C2 <- ggplot(pc_var_C2, aes(x = factor(PC), y = Variance)) +
    geom_bar(stat = "identity", fill = "gray") +
    labs(x = "Principal Component", y = "Percentage of Variance Explained (%)") +
    ggtitle("Scree Plot of PCA in C2") +
    theme_minimal()

# Print the barplot
barplot_C2

# Print Loadings Correlation Matrix
cor_loadings_C2 <- pca_results_C2$rotation  # PCA loadings
cor_matrix_C2 <- cor(cor_loadings_C2)
cor_matrix_C2

names(explained_variance_perc_C2) <- colnames(cor_matrix_C2)

```

## Export of figures and data – Core C2

### Figure export

PCA visualizations are exported in both vector and raster formats:

- **SVG format**: preserves vector quality for publication and editing (e.g., Illustrator)  
- **TIFF format (300 dpi)**: suitable for journal submission requirements  

Exported figures include:

- PCA loadings correlation biplot  
- PCA score plot faceted by facies  
- PCA score plot (combined facies)  
- Scree plot (explained variance)  

All figures are saved with standardized dimensions (6 × 6 inches) to ensure consistency across outputs.

### Data export

Processed datasets are exported as `.csv` files to facilitate interoperability with external workflows:

- `clr_data_C2.csv`: clr-transformed elemental dataset  
- `PCA_data_C2.csv`: PCA scores (PC1, PC2, depth, facies)  
- `C2_Variance_Perc.csv`: percentage of variance explained per component  
- `correlation_matrix_C2.csv`: correlation matrix derived from PCA loadings  

The clr and PCA score outputs are specifically intended for downstream visualization and stratigraphic plotting in MATLAB-based workflows.

### Workflow integration

Exporting intermediate and final datasets ensures:

- Reproducibility of results  
- Flexibility for cross-platform analysis (R ↔ MATLAB)  
- Efficient separation between statistical analysis and visualization pipelines  

### Workspace cleanup

Intermediate objects and large datasets are removed after export to:

- Reduce memory usage  
- Prevent object conflicts in subsequent analyses  
- Maintain a clean working environment  

### Notes

- File paths assume a predefined directory structure (`plots/C2/`, `outputFiles/C2/`)  
- Users should ensure directories exist prior to execution or implement automated directory creation  
- Retaining exported `.csv` files allows reproducibility without rerunning PCA computations  

```{r Save Figures and data for C2}

#Saving PCA plots
ggsave("plots/C2/pca_biplot_cor_C2.svg", pca_biplot_cor_C2, width = 6, height = 6, units = "in", device = "svg")
ggsave("plots/C2/pca_biplot_cor_C2.tiff", pca_biplot_cor_C2, width = 6, height = 6, dpi = 300)
ggsave("plots/C2/pcaC2_biplot_facies_facets_C2.svg", pcaC2_biplot_facies_facets_C2, width = 6, height = 6, units = "in", device = "svg")
ggsave("plots/C2/pcaC2_biplot_facies_facets_C2.tiff", pcaC2_biplot_facies_facets_C2, width = 6, height = 6, dpi = 300)

ggsave("plots/C2/pcaC2_biplot_facies.svg", pcaC2_biplot_facies, width = 6, height = 6, units = "in", device = "svg")


#Save barplot
ggsave("plots/C2/barplot_C2.svg", barplot_C2, width = 6, height = 6, units = "in", device = "svg")


#Save CLR and PCA scores for MATLAB plotting (I will use this for downcore plotting, which is the code in MATLAB that is not here)
write.csv(clr_data_C2, file = "outputFiles/C2/clr_data_C2.csv", row.names = FALSE)
write.csv(scores_C2, file = "outputFiles/C2/PCA_data_C2.csv", row.names = FALSE)
write.csv(explained_variance_perc_C2, file = "outputFiles/C2/C2_Variance_Perc.csv", row.names = FALSE)

#Save correlation Matrix as a CSV file

write.csv(cor_matrix_C2, file = "outputFiles/C2/correlation_matrix_C2.csv", row.names = T)

rm(pca_biplot_cor_C2)
rm(pcaC2_biplot_facies_facets_C2)
rm(cor_matrix_C2)
rm(scores_C2)
rm(XRF_C2)
rm(clr_data_C2)
rm(pca_results_C2)
rm(loadings_C2)
rm(pc_var_C2)
rm(cor_loadings_C2)
rm(barplot_C2)

```

# C3 Analysis
## Principal Component Analysis (PCA) – Core C3

The PCA workflow applied to Core C3 follows the same methodology described for Core C2, including:

- Subsetting clr-transformed data to remove non-compositional or artifact-prone variables  
- Performing PCA using centered and scaled variables  
- Extracting PCA scores (PC1 and PC2) and loadings  
- Visualizing results through score plots, faceted plots, and loadings (correlation biplots)  
- Calculating explained variance and generating scree plots  
- Exporting figures and datasets for downstream analysis  

### Facies classification

For Core C3, stratigraphic facies are defined using the following depth intervals:

- ZC3A: 0–15 cm  
- ZC3B: 15–38 cm  
- ZC3C: 38–53.6 cm  
- ZC3D: 53.6–94 cm  

These intervals reflect the vertical structure of the core and are used to interpret PCA results in a stratigraphic and paleoenvironmental context.

### Notes

- All preprocessing, transformation, and visualization steps are identical to those applied to Core C2  
- Differences between cores arise solely from the selected elements and stratigraphic zoning  
- Maintaining methodological consistency allows direct comparison of PCA results between cores  


```{r PCA C3}

# --- Subset Data for PCA ---
# Only include the elements you wish to use for the PCA
subset_data_C3 <- clr_data_C3 %>% dplyr::select(-c("Depth", "K", "Cl"))

# --- Perform PCA on clr-transformed data ---
pca_results_C3 <- prcomp(subset_data_C3, center = TRUE, scale. = TRUE)

# --- Extract PCA Scores ---
scores_C3 <- data.frame(pca_results_C3$x[, 1:2])
colnames(scores_C3) <- c("PC1", "PC2")

# --- Add Depth and Facies Information ---
# Here, we assume you have a depth vector (Depth_cm) from your original data.
# Adjust this accordingly if Depth_cm is stored elsewhere.
scores_C3$Depth <- clr_data_C3$Depth  # Replace 'original_data' with your depth data frame

# Create facies based on depth intervals:
# ZC3A: 55-95 cm, ZC3B: 40-55 cm, ZC3C: 20-40 cm, ZC2D: 0-20 cm.
scores_C3$Facies <- cut(scores_C3$Depth,
                     breaks = c(0, 15, 38, 53.6, 94),
                     labels = c("ZC3A", "ZC3B", "ZC3C", "ZC3D"),
                     include.lowest = TRUE, right = TRUE)

#Eliminate ant NaN values
scores_C3 <- scores_C3[complete.cases(scores_C3), ]


# --- Extract and scale Loadings ---
loadings_C3 <- pca_results_C3$rotation[, 1:2] * 
  matrix(pca_results_C3$sdev[1:2], nrow = nrow(pca_results_C3$rotation), ncol = 2, byrow = TRUE)
loadings_C3 <- data.frame(loadings_C3)
colnames(loadings_C3) <- c("PC1", "PC2")
loadings_C3$variable <- rownames(loadings_C3)

####Scale Loadings to the same coordinate system as the scores (using the maximum absolute score). OPTIONAL###
#scale_factor_C3 <- max(abs(scores_C3$PC1), abs(scores_C3$PC2))
#loadings_scaled_C3 <- loadings_C3
#loadings_scaled_C3$PC1 <- loadings_scaled_C3$PC1 * scale_factor_C3
#loadings_scaled_C3$PC2 <- loadings_scaled_C3$PC2 * scale_factor_C3

rm(subset_data_C3)

```

```{r Plot C3 Individual Factor Map}

# --- Create the PCA Biplot (Individual Factor Map) ---
pcaC3_biplot_facies <- ggplot() +
  # Plot sample scores, colored by facies
  geom_point(data = scores_C3, aes(x = PC1, y = PC2, color = Facies, text = Depth), alpha = 0.7, size = 2) +
  theme_minimal() +
  labs(title = "PCA Biplot with Facies for Core C3",
       x = "PC1", y = "PC2", color = "Facies")

# Print the plot
print(pcaC3_biplot_facies)

interactive_plot_PCA_C3 <- ggplotly(pcaC3_biplot_facies, tooltip = c("x", "y", "color", "text"))
interactive_plot_PCA_C3

#Plot separate as facets
pcaC2_biplot_facies_facets_C3 <- ggplot() +
  # Plot sample scores, colored by facies
  geom_point(data = scores_C3, aes(x = PC1, y = PC2, color = Facies), alpha = 0.7, size = 2) +
  facet_wrap(~ Facies) +
  theme_minimal() +
  labs(title = "PCA Biplot Faceted by Facies for Core C3",
       x = "PC1", y = "PC2", color = "Facies")

pcaC2_biplot_facies_facets_C3



```




```{r Plot C3 only the eigenvectors without the scores}

# Create the correlation biplot with fixed axes [-1, 1]
pca_biplot_cor_C3 <- ggplot() +
  geom_segment(data = loadings_C3, 
               aes(x = 0, y = 0, xend = PC1, yend = PC2), 
               arrow = arrow(length = unit(0.2, "cm")), 
               color = "black") +
  geom_text_repel(data = loadings_C3, 
                  aes(x = PC1, y = PC2, label = variable), 
                  color = "red", size = 4) +
  theme_minimal() +
  labs(title = "Correlation Biplot from clr-transformed XRF Data in C3", 
       x = "PC1", y = "PC2") +
  xlim(-1, 1) +
  ylim(-1, 1)

print(pca_biplot_cor_C3)


```



```{r Explained variance and Correlation Matrix for C3}
# Explained Variance
explained_variance_C3 <- pca_results_C3$sdev^2 / sum(pca_results_C3$sdev^2)
explained_variance_perc_C3 <- explained_variance_C3 * 100

# Print Explained Variance
print(explained_variance_perc_C3)

# Creating the barplot for Explained Variance
pc_var_C3 <- data.frame(PC = seq(along = explained_variance_perc_C3), Variance = explained_variance_perc_C3)
barplot_C3 <- ggplot(pc_var_C3, aes(x = factor(PC), y = Variance)) +
    geom_bar(stat = "identity", fill = "gray") +
    labs(x = "Principal Component", y = "Percentage of Variance Explained (%)") +
    ggtitle("Scree Plot of PCA in C3") +
    theme_minimal()

# Print the barplot
barplot_C3

# Print Loadings Correlation Matrix
cor_loadings_C3 <- pca_results_C3$rotation  # PCA loadings
cor_matrix_C3 <- cor(cor_loadings_C3)
cor_matrix_C3

names(explained_variance_perc_C3) <- colnames(cor_matrix_C3)
```

```{r Save Figures and data for C3}

#Saving PCA plots
ggsave("plots/C3/pca_biplot_cor_C3.svg", pca_biplot_cor_C3, width = 6, height = 6, units = "in", device = "svg")
ggsave("plots/C3/pca_biplot_cor_C3.tiff", pca_biplot_cor_C3, width = 6, height = 6, dpi = 300)
ggsave("plots/C3/pcaC2_biplot_facies_facets_C3.svg", pcaC2_biplot_facies_facets_C3, width = 6, height = 6, units = "in", device = "svg")
ggsave("plots/C3/pcaC2_biplot_facies_facets_C3.tiff", pcaC2_biplot_facies_facets_C3, width = 6, height = 6, dpi = 300)

ggsave("plots/C3/pcaC3_biplot_facies.svg", pcaC3_biplot_facies, width = 6, height = 6, units = "in", device = "svg")


#Save barplot
ggsave("plots/C3/barplot_C3.svg", barplot_C3, width = 6, height = 6, units = "in", device = "svg")


#Save CLR and PCA scores for MATLAB plotting (I will use this for downcore plotting, which is the code in MATLAB that is not here)
write.csv(clr_data_C3, file = "outputFiles/C3/clr_data_C3.csv", row.names = FALSE)
write.csv(scores_C3, file = "outputFiles/C3/PCA_data_C3.csv", row.names = FALSE)
write.csv(explained_variance_perc_C3, file = "outputFiles/C3/C3_Variance_Perc.csv", row.names = FALSE)

#Save correlation Matrix as a CSV file

write.csv(cor_matrix_C3, file = "outputFiles/C3/correlation_matrix_C3.csv", row.names = T)


```
