---
title: "Code 1: XRF clr-transformation"
author: "Bryan Rodriguez-Colon"
output: html_document
date: "2024-04-30"
---

## Overview

This script performs preprocessing and centered log-ratio (clr) transformation of XRF-derived elemental count data from two sediment cores (C2 and C3) collected in southwestern Puerto Rico. The workflow prepares compositional geochemical data for downstream multivariate analyses (e.g., PCA) by addressing sparsity, noise, and constant-sum constraints inherent to XRF datasets.

## Objectives

1.  Import XRF elemental count data and associated depth vectors\
2.  Apply quality control filtering to remove unreliable elements\
3.  Transform compositional data using clr transformation\
4.  Export processed datasets for modular downstream analysis

## Workflow structure

The script is organized into four main stages:

### 1. Function definition

A custom function (`clr_transform()`) is defined to perform compositional transformation, including: - Zero replacement\
- Removal of incomplete observations\
- Calculation of geometric mean\
- clr transformation of elemental data\
- Reintegration of depth information

### 2. Data import

XRF datasets and corresponding depth vectors are imported for: - Core C2 (PS-C2)\
- Core C3 (PC-C3)

Data were originally processed in MATLAB using workflows provided by the Sediment and Coastal Dynamics Laboratory (UMass Amherst), and exported for use in R.

### 3. Data cleaning and element selection

Elemental data are filtered using both automated and manual criteria:

-   Removal of elements with:
    -   High sparsity (\>80% zeros)\
    -   Low signal intensity (\<40 mean counts)
-   Manual exclusion of elements with known analytical issues:
    -   Sb, Zr, Ge
-   Core-specific adjustments are applied where geochemically justified (e.g., inclusion of S and K in C3)

This step ensures that only statistically robust and geochemically meaningful variables are retained.

### 4. clr transformation

Filtered datasets are transformed using the centered log-ratio approach (Aitchison, 1986), which:

-   Removes compositional closure effects\
-   Enables valid multivariate statistical analysis\
-   Standardizes data across samples

The transformation is applied independently to each core.

### 5. Data export

The resulting clr-transformed datasets are saved as a single `.RData` <file:%60RData/clr_XRF_cores.RData>\`

This enables a modular workflow where downstream analyses (e.g., PCA) can be conducted without repeating preprocessing steps.

## Output

The script produces:

-   `clr_data_C2`: clr-transformed dataset for Core C2\
-   `clr_data_C3`: clr-transformed dataset for Core C3

Each dataset includes: - clr-transformed elemental variables\
- Depth information for stratigraphic context

## Reproducibility notes

-   Element selection directly influences clr space and must remain consistent across analyses\
-   Zero-replacement strategy follows standard compositional data practices but may influence low-abundance elements\
-   MATLAB preprocessing steps are external to this workflow and are not included in the repository

## References

Sebastien Bertrand, Rik Tjallingii, Malin E Kylander, Bruno Wilhelm, Stephen J Roberts, et al., *Inorganic geochemistry of lake sediments: A review of analytical techniques and guidelines for data interpretation.* Earth-Science Reviews, 2024, 249, pp.104639.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

rm(list = ls())
#setwd('/Users/bryanrodriguez/Library/Mobile Documents/com~apple~CloudDocs/PhD-PROJECT/R-codes/PRGeobio_XR_C14_StableIsotopes_Codes/XRF_Codes')

```

Stop data displayed in scientific notation

```{r stop scipen}
options(scipen = 999)
```

```{r load packages, inclue = T, message = F}

library(ggplot2)      # plotting (all figures)
library(dplyr)        # data manipulation (select, filter, pipes)

library(compositions) # clr conceptual framework
library(zCompositions) # zero-replacement (even if not used directly, justified)

library(MASS)         # LDA (lda function)
library(ggrepel)      # label placement in biplots

library(plotly)       # interactive plots
```

# Functions

## `clr_trasnform()`

The function `clr_transform()` performs a centered log-ratio (clr) transformation on XRF elemental count data to account for the compositional nature of geochemical datasets.

### Inputs

-   `data`: dataframe containing XRF elemental counts\
-   `selected_elements`: vector of element names to include in the transformation\
-   `depth_data`: vector of depth values corresponding to each sample

### Pre-processing steps

1.  **Element subsetting**\
    Only the selected elements are retained from the input dataset to ensure that the clr transformation is applied to a consistent compositional subset.

2.  **Zero replacement**\
    Zeros are replaced with half of the minimum non-zero value within each element. This avoids undefined logarithmic operations during transformation while preserving relative differences between samples.

3.  **Removal of incomplete cases**\
    Rows containing missing values are excluded to ensure numerical stability in the transformation and consistency between geochemical and depth data.

### clr transformation

For each sample, the clr transformation is computed as:

clr(xᵢ) = ln(xᵢ / g(x))

where:

-   xᵢ is the concentration (or count) of element *i*\
-   g(x) is the geometric mean of all elements in that sample

The geometric mean is calculated as:

g(x) = exp(mean(log(x)))

This transformation removes the constant-sum constraint inherent to compositional data and allows for valid multivariate statistical analysis (Aitchison, 1986).

### Output

The function returns a dataframe containing:

-   clr-transformed elemental values\
-   a `Depth` column appended as the first column for stratigraphic reference

### Notes

-   The clr transformation assumes that the data are compositional and strictly positive after zero replacement\
-   The choice of elements included in the transformation influences the resulting clr space and should be consistent across analyses\
-   This implementation follows standard compositional data analysis approaches commonly applied to XRF datasets in sedimentology and paleoenvironmental studies

```{r clr transformation function}
clr_transform <- function(data, selected_elements, depth_data) {
  # Subset the data to include only the selected elements
  subset_data <- data[, selected_elements]
  
  # Ensure depth_data is a vector (if it's not already)
  depth_vec <- depth_data
  
  # Replace zeros with half of the minimum non-zero value in each column
  replaceZeros <- function(x) {
    nonzero <- x[x > 0 & !is.na(x)]
    if (length(nonzero) > 0) {
      x[x == 0] <- min(nonzero) / 2
    }
    return(x)
  }
  subset_data_clean <- as.data.frame(lapply(subset_data, replaceZeros))
  
  # Identify the rows with complete cases (no NA values)
  complete_idx <- complete.cases(subset_data_clean)
  
  # Subset the cleaned data and the depth vector using the same indices
  subset_data_clean <- subset_data_clean[complete_idx, ]
  depth_clean <- depth_vec[complete_idx]
  
  # Compute the geometric mean for each row
  geomMean <- apply(subset_data_clean, 1, function(x) exp(mean(log(x))))
  
  # Compute the clr transformation for each element in each row
  clr_data <- subset_data_clean  # initialize a data frame for clr values
  for (i in 1:nrow(subset_data_clean)) {
    clr_data[i, ] <- log(subset_data_clean[i, ] / geomMean[i])
  }
  
  # Add the depth information to the clr_data dataframe as the first column
  clr_data$Depth <- depth_clean
  clr_data <- clr_data[, c("Depth", setdiff(names(clr_data), "Depth"))]
  
  return(clr_data)
}

```

```{r To reset data without erasing functions}
rm(list = ls()[!sapply(ls(), function(x) is.function(get(x)))])
```

# Loading data

## Data import

XRF elemental count data and corresponding depth vectors are imported for two sediment cores:

-   **C2 core (PS-C2)**\
-   **C3 core (PC-C3)**

### Input files

For each core, two files are used:

1.  **Elemental data (CSV format)**\
    Contains XRF elemental counts for each sample.

2.  **Depth data (TXT format)**\
    Contains corresponding depth values for each measurement.

#### ***Note on Data provenance***

The XRF datasets used in this script were exported from MATLAB-based processing workflows used to reduce and calibrate raw XRF scanner data. These MATLAB codes are not included in this repository, as they were not developed by the author. They were provided by co-author Pedro Matos-Llavona, who obtained them from the Sediment and Coastal Dynamics Laboratory at the University of Massachusetts Amherst.

### Data structure

-   Each row in the XRF dataset corresponds to a single measurement\
-   Columns represent elemental counts\
-   Depth vectors are imported separately and later aligned with geochemical data during preprocessing

This separation ensures flexibility in handling depth information and allows consistent indexing during clr transformation and downstream analyses.

```{r Load data}
#PS-C2
XRF_C2 <- read.csv("input/XRF_data_C2.csv")
# C2 depths
C2_depth <- scan("input/XRF_data_C2_depth.txt", sep = ",", what = numeric(), quiet = TRUE)

#PC-C3
XRF_C3 <- read.csv("input/XRF_data_C3.csv")
# C3 depths
C3_depth <- scan("input/XRF_data_C3_depth.txt", sep = ",", what = numeric(), quiet = TRUE)

```

#clr-transformation for both sediment cores

## Element selection and data cleaning prior to clr transformation

Prior to applying the centered log-ratio (clr) transformation, elemental data were filtered to remove noise and ensure statistical robustness.

### Elemental dataset

An initial set of 32 elements was considered:

Mg, Al, Si, P, S, Cl, Ar, K, Ca, Ti, V, Cr, Mn, Fe, Ni, Cu, Zn, Ge, As, Br, Rb, Sr, Y, Zr, Sb, Cs, Ba, Ce, Pr, Nd, Sm, Pb

### Automated filtering criteria

Elements were retained based on the following thresholds:

-   **Zero proportion threshold**: ≤ 80% zeros\
-   **Minimum mean count threshold**: ≥ 40 counts

These criteria remove elements with: - Excessive sparsity (high proportion of zeros) - Low signal intensity (close to detection limits)

### Manual quality control

In addition to automated filtering, the following elements were **manually excluded**:

-   Sb (antimony)\
-   Zr (zirconium)\
-   Ge (germanium)

These elements showed inconsistent signal behavior during exploratory analysis and are interpreted as likely analytical artifacts (e.g., spectral overlap or detector-related noise).

### Core-specific adjustments

-   For **C3**, sulfur (S) and potassium (K) were reintroduced after filtering due to their geochemical relevance in that core.

### Output

Two cleaned element sets are generated:

-   `elements_keep_C2`
-   `elements_keep_C3`

These represent the final variables used for compositional transformation and multivariate analysis.

Temporary variables used during filtering are removed to maintain a clean workspace.

```{r Clean element list for noise prior to clr-tranformation}

# All elements and set thresholds
elements <- c("Mg", "Al", "Si", "P", "S", "Cl", "Ar", "K", "Ca", "Ti", "V", "Cr", "Mn", "Fe", "Ni", "Cu", "Zn", "Ge", "As", "Br", "Rb", "Sr", "Y", "Zr", "Sb", "Cs", "Ba", "Ce", "Pr", "Nd", "Sm", "Pb") #32 elements

zero_threshold <- 0.80
count_threshold <- 40

# Identify numeric (element) columns only
elements_C2 <- XRF_C2[elements] %>% dplyr::select(where(is.numeric))
elements_C3 <- XRF_C3[elements] %>% dplyr::select(where(is.numeric))


# Calculate stats in one step
stats_C2 <- data.frame(
  element    = names(elements_C2),
  prop_zero  = colMeans(elements_C2 == 0, na.rm=TRUE),
  mean_count = colMeans(elements_C2, na.rm=TRUE)
)

stats_C3 <- data.frame(
  element    = names(elements_C3),
  prop_zero  = colMeans(elements_C3 == 0, na.rm=TRUE),
  mean_count = colMeans(elements_C3, na.rm=TRUE)
)

# Filter based on your thresholds
elements_keep_C2 <- stats_C2 %>%
  dplyr::filter(prop_zero <= zero_threshold, mean_count >= count_threshold) %>%
  pull(element)

elements_keep_C2 <- elements_keep_C2[!elements_keep_C2 %in% c("Sb", "Zr", "Ge")] #Take out Sb and Zr, as this is an analyitical error


elements_keep_C3 <- stats_C3 %>%
  dplyr::filter(prop_zero <= zero_threshold, mean_count >= count_threshold) %>%
  pull(element)

elements_keep_C3 <- c(elements_keep_C3, "S", "K") #Include Sulfur and potassium for C3

elements_keep_C3 <- elements_keep_C3[!elements_keep_C3 %in% c("Sb", "Zr", "Ge")]


print("C2 elements to keep:")
print(elements_keep_C2)

print("C3 elements to keep:")
print(elements_keep_C3)

rm(stats_C2)
rm(stats_C3)
rm(elements_C2)
rm(elements_C3)

# clr-transofrmation:

clr_data_C2 <- clr_transform(XRF_C2, elements_keep_C2, C2_depth)

clr_data_C3 <- clr_transform(XRF_C3, elements_keep_C3, C3_depth)

```

# Saving dataframes

```{r}

# Create output directory if it does not exist
if(!dir.exists("Rdata")){
  dir.create("RData", recursive = TRUE)
}

# Save both clr datasets together
save(clr_data_C2, clr_data_C3,
     file = "RData/clr_XRF_cores.RData")

print("files are now ready for Code 2: XRF PCA analysis")

```
