---
title: "XRD code 2: Data Alignment"
author: "Bryan Rodriguez"
date: "2025-03-22"
output: 
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    number_sections: true
    theme: united
    df_print: paged
    
description: This code portion is for alignment of all the samples (the ones with Calcite and dolomite displacements)
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
setwd("/Users/bryanrodriguez/Library/Mobile Documents/com~apple~CloudDocs/PhD-PROJECT/R-codes/PRGeobio_XRD_Isotopes_Codes/XRD_Codes")
```


```{r load packages, include=F, ECHO =F}
library(powdR)
library(ggplot2)
library(ggsci)
library(RColorBrewer)
library(plotly)
library(dplyr)
library(tibble)
library(purrr)
library(ggpubr)
library(lubridate)
library(patchwork)
library(baseline)
library(EEM)
library(data.table)
library(tidyverse)
library(conflicted)
library(stringr)
library(signal)
```


# Functions

## calculate_tth_AlignmentDiff function

The calculate_tth_AlignmentDiff function is designed to determine the difference in 2θ (tth) positions between two XRD spectra over a specified range. This is useful for aligning a sample spectrum to a standard by comparing their peak positions. The function takes two data frames (one for the sample and one for the standard) and a tth range. It then:

Subsets the Data:
Filters each data frame to only include rows where tth values fall within the specified range (tth_start to tth_end).
Identifies the Peak:
Finds the row with the highest count (intensity) in each subset. This row is assumed to represent the most prominent peak in that tth range.
Extracts the Peak Positions:
Retrieves the tth value corresponding to the highest intensity from both the sample and the standard.
Calculates the Difference:
Computes the difference between the sample’s and the standard’s peak positions. This difference can be used to adjust (or "shift") the sample spectrum for alignment with the standard.

```{r Alignment Function}

calculate_tth_AlignmentDiff <- function(df1, df2, tth_start, tth_end) {
  # Step 1: Subset the dataframes for tth values within the specified range
  subset_df1 <- df1[df1$tth >= tth_start & df1$tth <= tth_end, ]
  subset_df2 <- df2[df2$tth >= tth_start & df2$tth <= tth_end, ]
  
  # Step 2: Find the row with the highest count in each subset
  row_with_max_count_df1 <- subset_df1[which.max(subset_df1$counts), ]
  row_with_max_count_df2 <- subset_df2[which.max(subset_df2$counts), ]
  
  # Step 3: Extract the corresponding tth values
  tth_max_count_df1 <- row_with_max_count_df1$tth
  tth_max_count_df2 <- row_with_max_count_df2$tth
  
  # Step 4: Calculate the difference
  difference <- tth_max_count_df1 - tth_max_count_df2
  
  return(difference)
} #df1 is sample and df2 is Std. 
```


## Helper Function: align_sample

The align_sample function streamlines the alignment of XRD spectra by combining the calculation of the tth shift and the application of that shift in one step. It takes a sample spectrum and a standard spectrum along with a specified tth range (to calculate the shift) and an xlim range (to use in the alignment). If no manual shift is provided, it computes the shift using your existing calculate_tth_AlignmentDiff function, then applies the shift via align_xy. This modular approach reduces code repetition and makes it easier to adjust parameters for different samples.

```{r align_sample Function}
align_sample <- function(sample, standard, tth_range, xlim_range, manual_shift = NULL) {
  .Deprecated("align_sample_regrid")
  # If a manual shift is not provided, calculate it using the alignment function
  if (is.null(manual_shift)) {
    xshift <- calculate_tth_AlignmentDiff(sample, standard, tth_range[1], tth_range[2])
  } else {
    xshift <- manual_shift
  }
  # Align the sample using align_xy with the computed (or manual) shift
  aligned_sample <- align_xy(sample, std = standard, xmin = xlim_range[1], xmax = xlim_range[2], xshift = xshift)
  return(aligned_sample)
}

```

## New Align code to fix the the 2theta range differences after alignment
### Function: regrid_spectrum
Purpose:
The regrid_spectrum function re-interpolates an aligned XRD spectrum onto a common tth grid. This ensures that, regardless of any shifts or minor differences in the original data, every spectrum is represented by the same number of points (and identical tth values). This is particularly useful for comparing or combining spectra after alignment.

How It Works:

Define a Common Grid:
The function creates a sequence (common_grid) spanning from the lower to upper xlim limits, with a fixed number of points (n_points). For example, if you want exactly 3216 observations from 4.304275 to 59.15638, it generates a sequence with those properties.
Interpolation:
The function uses the approx() function to perform linear interpolation on the original spectrum. It takes the existing tth values and corresponding counts, and calculates the interpolated counts at each point on the common grid. The argument rule = 2 ensures that the boundary values are used for extrapolation if needed.
Output:
The function returns a new data frame containing the common tth grid and the interpolated counts. Every spectrum processed with this function will have the same tth values and the same number of observations.


```{r New functions to fix 2theta ranges}
regrid_spectrum <- function(spectrum, xlim_range, n_points = 3216) {
  common_grid <- seq(xlim_range[1], xlim_range[2], length.out = n_points)
  new_counts <- approx(x = spectrum$tth, y = spectrum$counts, 
                       xout = common_grid, rule = 2)$y
  return(data.frame(tth = common_grid, counts = new_counts))
}

```

### Function: align_sample_regrid

Purpose:
The align_sample_regrid function is an enhanced alignment helper. It not only calculates and applies the alignment shift (using calculate_tth_AlignmentDiff and align_xy), but also re-interpolates the aligned spectrum onto a common tth grid using the regrid_spectrum function. This ensures that every aligned spectrum ends up with the same fixed number of observations and tth values, which is essential for consistent downstream analysis.

How It Works:

Shift Calculation:
The function first checks if a manual shift has been provided via the manual_shift argument. By default, manual_shift is set to NULL, meaning no manual shift is provided. In that case, the function automatically calculates the alignment shift (xshift) by calling calculate_tth_AlignmentDiff over a specified tth range.
Manual Shift: If you supply a numeric value to manual_shift, the function will bypass the automatic calculation and use your provided value for the shift. This is useful if the automatic method does not yield satisfactory results for a particular sample.
Alignment:
The function applies the computed (or manual) shift to the sample using align_xy, which adjusts the sample’s tth values to align the key peak with the standard.
Regridding:
After alignment, the function calls regrid_spectrum to interpolate the aligned spectrum onto a common grid defined by xlim_range and with n_points observations. This step ensures that even if the alignment process changes the number of data points, every spectrum is re-interpolated to a consistent grid.
Output:
The function returns the regridded (aligned) spectrum as a data frame with standardized tth values and counts.

```{r New functions to fix 2theta ranges}
# Modified helper function that aligns a sample and then regrids it
align_sample_regrid <- function(sample, standard, tth_range, xlim_range, manual_shift = NULL, n_points = 3216) {
  # Compute the shift: automatically calculate if manual_shift is not provided
  if (is.null(manual_shift)) {
    xshift <- calculate_tth_AlignmentDiff(sample, standard, tth_range[1], tth_range[2])
  } else {
    xshift <- manual_shift
  }
  # Align the sample using align_xy with the computed (or manual) shift
  aligned_sample <- align_xy(sample, std = standard, xmin = xlim_range[1], xmax = xlim_range[2], xshift = xshift)
  # Re-interpolate the aligned sample onto the common tth grid
  regridded_sample <- regrid_spectrum(aligned_sample, xlim_range, n_points)
  return(regridded_sample)
}


```



# Load data

```{r To reset data without erasing functions}
rm(list = ls()[!sapply(ls(), function(x) is.function(get(x)))])
```


```{r Load data processed in code 1}

minstds_file_list <- readRDS("RData/min_standards.RDS")
MinStds2797_NoBkg <- minstds_file_list$MinStds2797_NoBkg_Smooth
rm(minstds_file_list)

DataNobkgSmooth_file_list <- readRDS("RData/Data_NoBkg_Smooth.RDS")

PL_NoBkg <- DataNobkgSmooth_file_list$PL_NoBkg_Smooth
PS_NoBkg <- DataNobkgSmooth_file_list$PS_NoBkg_Smooth
C2_NoBkg <- DataNobkgSmooth_file_list$C2_NoBkg_Smooth
C3_NoBkg <- DataNobkgSmooth_file_list$C3_NoBkg_Smooth
POC_NoBkg <- DataNobkgSmooth_file_list$POC_NoBkg_Smooth
POE_NoBkg <- DataNobkgSmooth_file_list$POE_NoBkg_Smooth
JB_NoBkg <- DataNobkgSmooth_file_list$JB_NoBkg
rm(DataNobkgSmooth_file_list)


```



# Alignment

The alignment process is critical for ensuring that key diffraction peaks in each spectrum line up with those of a reference standard. The following components are used:
Peak Difference Calculation: The calculate_tth_AlignmentDiff function identifies the peak in a specified tth range for both a sample and a standard, then calculates the difference (xshift) between their peak positions.
Alignment Helper Function: The custom align_sample function wraps the above difference calculation together with the application of the shift via align_xy. This modular approach reduces repetitive code and provides the flexibility to apply either automatic or manual shifts as needed.
Group-Specific Alignment: The code is organized into sections (e.g., Ponce Limestone, beach sands, microbialites, sediment cores) where each group of samples is aligned by comparing its spectra to a reference standard. For some samples, such as those with problematic automatic calculations (e.g., PL-5), a manual xshift is provided.

## Initial Standard Alingments

```{r  Standard Alignment}
# Align Aragonite_Halite_coral_standard using NewHalite_standard_S2797:
MinStds2797_NoBkg$Aragonite_Halite_coral_standard_S2797 <- align_sample_regrid(
  sample    = MinStds2797_NoBkg$Aragonite_Halite_coral_standard_S2797,
  standard  = MinStds2797_NoBkg$NewHalite_standard_S2797,
  tth_range = c(31.7, 32.07),            # Range to calculate xshift using Halite peak
  xlim_range = c(4.304275, 74.8),        # Range used in align_xy
  n_points = 2797
)

# Then, align Aragonite_coral_standard using the already aligned Aragonite_Halite_coral_standard:
MinStds2797_NoBkg$Aragonite_coral_standard_S2797 <- align_sample_regrid(
  sample    = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  standard  = MinStds2797_NoBkg$Aragonite_Halite_coral_standard_S2797,
  tth_range = c(26.04, 26.6),
  xlim_range = c(4.304275, 74.8),
  n_points = 2797
)

print("Standard Alignment done")
```

## Beach sand alingments

```{r Beach Sand Alignment}
# Parametes for all the beach samples
tth_range_BeachSand <- c(26.04, 26.6)
xlim_range_BeachSand <- c(4.304275, 74.8)

### For PO Sand:####
PL_NoBkg$Sand_PO2023 <- align_sample_regrid(
  sample    = PL_NoBkg$Sand_PO2023,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_BeachSand,
  xlim_range = xlim_range_BeachSand,
  n_points = 2797
)

### For JB Sand, using the Ooids standard:####
PL_NoBkg$Sand_JungleBeach <- align_sample_regrid(
  sample    = PL_NoBkg$Sand_JungleBeach,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_BeachSand,
  xlim_range = xlim_range_BeachSand,
  n_points = 2797
)

### For Canal Sand:####
PL_NoBkg$Sand_Canal <- align_sample_regrid(
  sample    = PL_NoBkg$Sand_Canal,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_BeachSand,
  xlim_range = xlim_range_BeachSand,
  n_points = 2797
)

print("Beach Sand Alignment done")
```

## Microbial Deposit Alignments

I created the plot below to double check the XRD peaks of the microbial deposits prior to alignment
```{r Checking plots before alingnment (code 2)}
#Quality Check Plotting
List_JB <- as_multi_xy(list( "JB19_Crust.green" = JB_NoBkg$JB19_Crust.green, "JB19_D.pink" = JB_NoBkg$JB19_D.pink, "JB19_Center.black" = JB_NoBkg$JB19_Center.black, "gypsum" = MinStds2797_NoBkg$Gypsum_S2797, "aragonite" = MinStds2797_NoBkg$Aragonite_coral_standard_S2797, "halite" = MinStds2797_NoBkg$Halite_S2797, "calcite" = MinStds2797_NoBkg$Calcite_S2797, "HMC" = MinStds2797_NoBkg$HigMgCal_S2797, "dolomite" = MinStds2797_NoBkg$Dolomite_S2797))
                             
List_POE <- as_multi_xy(list("PO19_01E01_crust.1" = POE_NoBkg$PO19_01E01_crust.1, "PO19_01E01_green.2" = POE_NoBkg$PO19_01E01_green.2, "PO19_01E01_brown.3" = POE_NoBkg$PO19_01E01_brown.3, "PO19_01E01_black.4" = POE_NoBkg$PO19_01E01_black.4, "aragonite" = MinStds2797_NoBkg$Aragonite_coral_standard_S2797, "halite" = MinStds2797_NoBkg$Halite_S2797, "calcite" = MinStds2797_NoBkg$Calcite_S2797, "HMC" = MinStds2797_NoBkg$HigMgCal_S2797, "dolomite" = MinStds2797_NoBkg$Dolomite_S2797))

List_POC <- as_multi_xy(list("PO19_01C03_Crust_green" = POC_NoBkg$PO19_01C03_Crust_green, "PO19_01C03_Green" = POC_NoBkg$PO19_01C03_Green, "PO19_01C03_grains" = POC_NoBkg$PO19_01C03_grains, "PO19_01C03_Black" = POC_NoBkg$PO19_01C03_Black, "PO19_01C03_Bottom_sand" = POC_NoBkg$PO19_01C03_Bottom_sand, "aragonite" = MinStds2797_NoBkg$Aragonite_coral_standard_S2797, "halite" = MinStds2797_NoBkg$Halite_S2797, "calcite" = MinStds2797_NoBkg$Calcite_S2797, "HMC" = MinStds2797_NoBkg$HigMgCal_S2797, "dolomite" = MinStds2797_NoBkg$Dolomite_S2797))

List_PS19_New <- as_multi_xy(list("PS19_Upper.crust1" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust1, "PS19_Upper.crust2" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust2, "PS19_Middle.Lithifed" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Middle.Lithifed, "PS19_Black.Layer" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Black.Layer, "PS19_L.below.Black" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_L.below.Black, "PS19_Bottom.Lithified" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Bottom.Lithified, "aragonite" = MinStds2797_NoBkg$Aragonite_coral_standard_S2797, "halite" = MinStds2797_NoBkg$Halite_S2797, "calcite" = MinStds2797_NoBkg$Calcite_S2797, "HMC" = MinStds2797_NoBkg$HigMgCal_S2797, "dolomite" = MinStds2797_NoBkg$Dolomite_S2797))

plot(List_POE, wavelength = "Cu", normalise = T, interactive = T)
```


```{r Microbial Deposits Alignment}
## PO Samples####
### For PO-01C3 (2019):


#### PO-01C3 Crust/green
POC_NoBkg$PO19_01C03_Crust_green <- align_sample_regrid(
  sample    = POC_NoBkg$PO19_01C03_Crust_green,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(8.6, 59.2),
  n_points = 2969
)

#### PO-01C3 Green
POC_NoBkg$PO19_01C03_Green <- align_sample_regrid(
  sample    = POC_NoBkg$PO19_01C03_Green,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(8.6, 59.2),
  n_points = 2969
)

#### PO-01C3 Grains
POC_NoBkg$PO19_01C03_grains <- align_sample_regrid(
  sample    = POC_NoBkg$PO19_01C03_grains,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(8.6, 59.2),
  n_points = 2969
)

#### PO-01C3 Black Layer
POC_NoBkg$PO19_01C03_Black <- align_sample_regrid(
  sample    = POC_NoBkg$PO19_01C03_Black,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(8.6, 59.2),
  n_points = 2969
)

#### PO-01C3 Black Layer
POC_NoBkg$PO19_01C03_Bottom_sand <- align_sample_regrid(
  sample    = POC_NoBkg$PO19_01C03_Bottom_sand,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(8.6, 59.2),
  n_points = 2969
)



### For PO-01E1 (2019):

#### Crust
POE_NoBkg$PO19_01E01_crust.1 <- align_sample_regrid(
  sample    = POE_NoBkg$PO19_01E01_crust.1,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(8.6, 59.2),
  n_points = 2971
)

#### Green
POE_NoBkg$PO19_01E01_green.2 <- align_sample_regrid(
  sample    = POE_NoBkg$PO19_01E01_green.2,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(8.6, 59.2),
  n_points = 2971
)

#### Brown
POE_NoBkg$PO19_01E01_brown.3 <- align_sample_regrid(
  sample    = POE_NoBkg$PO19_01E01_brown.3,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(8.6, 59.2),
  n_points = 2971
)

#### Black Layer
POE_NoBkg$PO19_01E01_black.4 <- align_sample_regrid(
  sample    = POE_NoBkg$PO19_01E01_black.4,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(8.6, 59.2),
  n_points = 2971
)



## PS19 Samples####

# For PS19 (Crust 1):
PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust1 <- align_sample_regrid(
  sample    = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust1,
  standard  = MinStds2797_NoBkg$Halite_S2797,
  tth_range = c(31.7, 32.07),
  xlim_range = c(4.304275, 59.15638),
  n_points = 3216
)

# For PS19 (Crust 2):
PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust2 <- align_sample_regrid(
  sample    = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust2,
  standard  = MinStds2797_NoBkg$Halite_S2797,
  tth_range = c(31.7, 32.07),
  xlim_range = c(4.304275, 59.15638),
  n_points = 3216
)


# For PS19 (Middle Lithified Layer):
PS_NoBkg$PS19_NoBkg_Smooth$PS19_Middle.Lithifed <- align_sample_regrid(
  sample    = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Middle.Lithifed,
  standard  = MinStds2797_NoBkg$Halite_S2797,
  tth_range = c(31.7, 32.07),
  xlim_range = c(4.304275, 59.15638),
  n_points = 3216
)


# For PS19 (layer with disordered Ca-dolomite):
PS_NoBkg$PS19_NoBkg_Smooth$PS19_L.below.Black <- align_sample_regrid(
  sample    = PS_NoBkg$PS19_NoBkg_Smooth$PS19_L.below.Black,
  standard  = MinStds2797_NoBkg$Halite_S2797,
  tth_range = c(31.7, 32.07),
  xlim_range = c(4.304275, 59.15638),
  n_points = 3216
)

# For PS19 (Black layer):
PS_NoBkg$PS19_NoBkg_Smooth$PS19_Black.Layer <- align_sample_regrid(
  sample    = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Black.Layer,
  standard  = MinStds2797_NoBkg$Halite_S2797,
  tth_range = c(31.7, 32.07),
  xlim_range = c(4.304275, 59.15638),
  n_points = 3216
)

# For PS19 (Bottom layer):
PS_NoBkg$PS19_NoBkg_Smooth$PS19_Bottom.Lithified <- align_sample_regrid(
  sample    = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Bottom.Lithified,
  standard  = MinStds2797_NoBkg$Halite_S2797,
  tth_range = c(31.7, 32.07),
  xlim_range = c(4.304275, 59.15638),
  n_points = 3216
)

## PS23 Samples####
# For PS23 (2023 microbialite):
PS_NoBkg$PS2223_NoBkg_Smooth$PS23_Bottom <- align_sample_regrid(
  sample    = PS_NoBkg$PS2223_NoBkg_Smooth$PS23_Bottom,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(4.31, 74.8),
  n_points = 2797
)


## PS Transects (2022) Samples####

### PST002#####
#PST-002 Crust
PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_002_Sam1_cbb <- align_sample_regrid(
  sample    = PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_002_Sam1_cbb,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(4.31, 74.8),
  n_points = 2797
)

#PST-002 Middle Sediment
PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_002_Sam2bsed <- align_sample_regrid(
  sample    = PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_002_Sam2bsed,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(4.31, 74.8),
  n_points = 2797
)

#PST-002 Middle Sediment 2
PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_002_Sam2dol <- align_sample_regrid(
  sample    = PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_002_Sam2dol,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(4.31, 74.8),
  n_points = 2797
)

#PST-002 Bottom Sediment
PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_002_Sam3bottom <- align_sample_regrid(
  sample    = PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_002_Sam3bottom,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(4.31, 74.8),
  n_points = 2797
)

### PST005#####
#PST-005 Crust & middle
PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_005_Sam1CrustMiddle <- align_sample_regrid(
  sample    = PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_005_Sam1CrustMiddle,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(4.31, 74.8),
  n_points = 2797
)

#PST-005 Bottom
PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_005_Sam2Bottom <- align_sample_regrid(
  sample    = PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_005_Sam2Bottom,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = c(4.31, 74.8),
  n_points = 2797
)

### PS-Farm#####
#PST- Farm
PS_NoBkg$PS2223_NoBkg_Smooth$PS22_Farm <- align_sample_regrid(
  sample    = PS_NoBkg$PS2223_NoBkg_Smooth$PS22_Farm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.05, 26.7),
  xlim_range = c(4.31, 74.8),
  n_points = 2797
)


print("Microbial Deposits Alignment done")

```

## Ponce Limestone Alignment

The Ponce Limestone samples have properties that cause the automatic calculation of the alignment shift (via calculate_tth_AlignmentDiff) to fail or produce unsatisfactory results. In those cases, the custom helper function (align_sample) may not work directly without manual intervention. For example, with PL-5, the min and max functions on the tth values did not yield the appropriate range for calculating the shift, so a manual xshift value (0.09213) was used instead.

```{r Ponce Limestone Alignment}
### Ponce Limestone Alignment

## PL-13:####
#Align the Halite-containing version first using the NewHalite standard####
tth_diff13Hal <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_13_Halite, 
                                               MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                               31.6, 32.2)
PL_NoBkg$PL_13_Halite <- align_xy(PL_NoBkg$PL_13_Halite, 
                                  std = MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                  xmin = 4.304275, xmax = 74.8, 
                                  xshift = tth_diff13Hal)

tth_diff13Final <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_15_NoHalite, 
                                                PL_NoBkg$PL_15_Halite, 
                                                30.7, 31.3)
PL_NoBkg$PL_13_NoHalite <- align_xy(PL_NoBkg$PL_13_NoHalite, 
                                    std = PL_NoBkg$PL_13_Halite, 
                                    xmin = min(PL_NoBkg$PL_13_Halite$tth), 
                                    xmax = max(PL_NoBkg$PL_13_Halite$tth), 
                                    xshift = tth_diff13Final)

## PL-15:####
#First, align the Halite version using the NewHalite standard, then use it to align the NoHalite version.
tth_diff15Hal <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_15_Halite, 
                                              MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                              31.7, 32.4)
PL_NoBkg$PL_15_Halite <- align_xy(PL_NoBkg$PL_15_Halite, 
                                  std = MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                  xmin = 4.304275, xmax = 74.8, 
                                  xshift = tth_diff15Hal)

tth_diff15Final <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_15_NoHalite, 
                                                PL_NoBkg$PL_15_Halite, 
                                                30.7, 31.3)
PL_NoBkg$PL_15_NoHalite <- align_xy(PL_NoBkg$PL_15_NoHalite, 
                                    std = PL_NoBkg$PL_15_Halite, 
                                    xmin = min(PL_NoBkg$PL_15_Halite$tth), 
                                    xmax = max(PL_NoBkg$PL_15_Halite$tth), 
                                    xshift = tth_diff15Final)

## PL-5:####
#Align the WithHalite version, then use a manual shift for the NoHalite version
tth_diff5Hal <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_5_WithHalite, 
                                             MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                             31.6, 32.05)
PL_NoBkg$PL_5_WithHalite <- align_xy(PL_NoBkg$PL_5_WithHalite, 
                                      std = MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                      xmin = 4.304275, xmax = 74.8, 
                                      xshift = tth_diff5Hal)
# Here the automatic function did not work properly, so we use a manual shift (0.09213)
PL_NoBkg$PL_5_NoHalite <- align_xy(PL_NoBkg$PL_5_NoHalite, 
                                   std = PL_NoBkg$PL_5_WithHalite, 
                                   xmin = min(PL_NoBkg$PL_5_WithHalite$tth), 
                                   xmax = max(PL_NoBkg$PL_5_WithHalite$tth), 
                                   xshift = 0.09213)

## PL-6:####
#Align the Halite version then use it to align the NoHalite version
tth_diff6Hal <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_6_Halite, 
                                             MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                             31.6, 32.05)
PL_NoBkg$PL_6_Halite <- align_xy(PL_NoBkg$PL_6_Halite, 
                                  std = MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                  xmin = 4.304275, xmax = 74.8, 
                                  xshift = tth_diff6Hal)

tth_diff6Final <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_6_NoHalite, 
                                               PL_NoBkg$PL_6_Halite, 
                                               30.9, 31.2)
PL_NoBkg$PL_6_NoHalite <- align_xy(PL_NoBkg$PL_6_NoHalite, 
                                   std = PL_NoBkg$PL_6_Halite, 
                                   xmin = min(PL_NoBkg$PL_6_Halite$tth), xmax = 74, 
                                   xshift = tth_diff6Final)

## PL-9:#### 
#Align the Halite version then align the NoHalite version
tth_diff9Hal <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_9_Halite, 
                                             MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                             31.3, 32.5)
PL_NoBkg$PL_9_Halite <- align_xy(PL_NoBkg$PL_9_Halite, 
                                  std = MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                  xmin = 4.304275, xmax = 74.8, 
                                  xshift = tth_diff9Hal)
tth_diff9Final <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_9_NoHalite, 
                                               PL_NoBkg$PL_9_Halite, 
                                               29.06, 30.04)
PL_NoBkg$PL_9_NoHalite <- align_xy(PL_NoBkg$PL_9_NoHalite, 
                                   std = PL_NoBkg$PL_9_Halite, 
                                   xmin = min(PL_NoBkg$PL_9_Halite$tth), 
                                   xmax = max(PL_NoBkg$PL_9_Halite$tth), 
                                   xshift = tth_diff9Final)

## PL-1:####
#Align the Halite version then the NoHalite version
tth_diff1Hal <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_1_Halite, 
                                             MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                             31.3, 32.5)
PL_NoBkg$PL_1_Halite <- align_xy(PL_NoBkg$PL_1_Halite, 
                                  std = MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                  xmin = 4.304275, xmax = 74.8, 
                                  xshift = tth_diff1Hal)
tth_diff1Final <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_1_NoHalite, 
                                               PL_NoBkg$PL_1_Halite, 
                                               29.06, 30.04)
PL_NoBkg$PL_1_NoHalite <- align_xy(PL_NoBkg$PL_1_NoHalite, 
                                   std = PL_NoBkg$PL_1_Halite, 
                                   xmin = min(PL_NoBkg$PL_1_Halite$tth), 
                                   xmax = max(PL_NoBkg$PL_1_Halite$tth), 
                                   xshift = tth_diff1Final)

## PL-8:####
#Align the Halite version then the NoHalite version
tth_diff8Hal <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_8_Halite, 
                                             MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                             31.3, 32.5)
PL_NoBkg$PL_8_Halite <- align_xy(PL_NoBkg$PL_8_Halite, 
                                  std = MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                  xmin = 4.304275, xmax = 74.8, 
                                  xshift = tth_diff8Hal)
tth_diff8Final <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_8_NoHalite, 
                                               PL_NoBkg$PL_8_Halite, 
                                               29.06, 30.04)
PL_NoBkg$PL_8_NoHalite <- align_xy(PL_NoBkg$PL_8_NoHalite, 
                                   std = PL_NoBkg$PL_8_Halite, 
                                   xmin = min(PL_NoBkg$PL_8_Halite$tth), 
                                   xmax = max(PL_NoBkg$PL_8_Halite$tth), 
                                   xshift = tth_diff8Final)

## PL-Canal:####
#Align the Halite version then use it to align the canal sample
tth_diffCanalHal <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_canal_Halite, 
                                                 MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                                 31.3, 32.5)
PL_NoBkg$PL_canal_Halite <- align_xy(PL_NoBkg$PL_canal_Halite, 
                                      std = MinStds2797_NoBkg$NewHalite_standard_S2797, 
                                      xmin = 4.304275, xmax = 74.8, 
                                      xshift = tth_diffCanalHal)
tth_diffCanalFinal <- calculate_tth_AlignmentDiff(PL_NoBkg$PL_Canal, 
                                                   PL_NoBkg$PL_canal_Halite, 
                                                   29.06, 30.04)
PL_NoBkg$PL_Canal <- align_xy(PL_NoBkg$PL_Canal, 
                             std = PL_NoBkg$PL_canal_Halite, 
                             xmin = min(PL_NoBkg$PL_canal_Halite$tth), xmax = 74, 
                             xshift = tth_diffCanalFinal)


print("Ponce Limestone Alignment done")
```

```{r Eliminate all values to understature the Environment}
rm(list = ls()[sapply(ls(), function(x) {
  obj <- get(x)
  is.atomic(obj) || is.factor(obj)
})])

```



## Sediment Cores Alignments


```{r parameters for Sediment core samples alignments}
### Sediment Cores Alignment

# Define common parameters for the alignment of the sediment cores
tth_range_core <- c(26.1, 26.5)    # tth range over which to calculate the shift. Only C3 52cm was put manually.
xlim_core      <- c(4.31, 74)       # xlim range to use in align_xy
C2_npoints <- 8389
```

```{r C2 core}

# Align each depth in the C2 core using the same standard
C2_NoBkg$d112cm <- align_sample_regrid(
  sample    = C2_NoBkg$d112cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = C2_npoints
)

C2_NoBkg$d113cm <- align_sample_regrid(
  sample    = C2_NoBkg$d113cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = C2_npoints
)

C2_NoBkg$d103cm <- align_sample_regrid(
  sample    = C2_NoBkg$d103cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = C2_npoints
)

C2_NoBkg$d115cm <- align_sample_regrid(
  sample    = C2_NoBkg$d115cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = C2_npoints
)

C2_NoBkg$d46cm <- align_sample_regrid(
  sample    = C2_NoBkg$d46cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = C2_npoints
)

C2_NoBkg$d56cm <- align_sample_regrid(
  sample    = C2_NoBkg$d56cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = C2_npoints
)

C2_NoBkg$d58cm <- align_sample_regrid(
  sample    = C2_NoBkg$d58cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = C2_npoints
)

C2_NoBkg$d74cm <- align_sample_regrid(
  sample    = C2_NoBkg$d74cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = C2_npoints
)

C2_NoBkg$d94cm <- align_sample_regrid(
  sample    = C2_NoBkg$d94cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = C2_npoints
)

print("C2 Alignment done")
```

```{r C3 Core}


## --- For Core C3 ---
# Align each depth in the C3 core using the same standard
C3_NoBkg$d49_gs_cm <- align_sample_regrid(
  sample    = C3_NoBkg$d49_gs_cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = 2797
)

C3_NoBkg$d52cm <- align_sample_regrid(
  sample    = C3_NoBkg$d52cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = c(26.1, 26.6),
  xlim_range = xlim_core,
  n_points = 2797
)

C3_NoBkg$d44cm <- align_sample_regrid(
  sample    = C3_NoBkg$d44cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = 2797
)

C3_NoBkg$d47cm <- align_sample_regrid(
  sample    = C3_NoBkg$d47cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = 2797
)

C3_NoBkg$d42cm <- align_sample_regrid(
  sample    = C3_NoBkg$d42cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = 2797
)

C3_NoBkg$d39cm <- align_sample_regrid(
  sample    = C3_NoBkg$d39cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = 2797
)

C3_NoBkg$d37cm <- align_sample_regrid(
  sample    = C3_NoBkg$d37cm,
  standard  = MinStds2797_NoBkg$Aragonite_coral_standard_S2797,
  tth_range = tth_range_core,
  xlim_range = xlim_core,
  n_points = 2797
)

print("C3 Alignment done")

```




# Saving Corrected Data

```{r Saving Data}

Aligend_Data_NoBkg <- list(PL_NoBkg_Aligned_Smooth = PL_NoBkg, C3_NoBkg_Aligned_Smooth = C3_NoBkg, C2_NoBkg_Aligned_Smooth = C2_NoBkg, PS_NoBkg_Aligned_Smooth = PS_NoBkg, POC_NoBkg_Aligned_Smooth = POC_NoBkg, POE_NoBkg_Aligned_Smooth = POE_NoBkg, JB_NoBkg_Aligned_Smooth = JB_NoBkg)


#Save 
saveRDS(Aligend_Data_NoBkg, file = "RData/Aligned_Data_NoBkg_Smooth.RDS")


print("All data aligned and saved")
```






