---
title: 'XRD Code 3: Plotting'
author: "Bryan Rodriguez"
date: "2025-03-25"
output: 
  html_document:
    self_contained: true
    toc: true
    toc_float: true
    number_sections: true
    theme: united
    df_print: paged
Description: This code portion is only to plot the XRD data
---



```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
setwd("~/OneDrive - University of Kansas/Bryan-MS_PhD/PHD-PROJECT/PuertoRico/CHAPTER3/XRD/R_plotting")

```


```{r load packages, include=F, ECHO =F}
library(powdR)
library(ggplot2)
library(ggsci)
library(RColorBrewer)
library(plotly)
library(dplyr)
library(tibble)
library(purrr)
library(ggpubr)
library(lubridate)
library(patchwork)
library(baseline)
library(EEM)
library(data.table)
library(tidyverse)
library(conflicted)
library(stringr)
library(grid) 
library(ggh4x) 
library(htmlwidgets)
```

# Load data

```{r Data}
#Standards
minstds_file_list <- readRDS("RData/min_standards.RDS")
MinStds8389_NoBkg <-minstds_file_list$MinStds8389_NoBkg_Smooth
MinStds2797_NoBkg <- minstds_file_list$MinStds2797_NoBkg_Smooth
MinStds2797_NoBkg <- as_multi_xy(MinStds2797_NoBkg)
rm(minstds_file_list)

# Main peaks
mins_mainPeaks <- readRDS("RData/mins_mainPeaks.RDS")

# Data
DataNobkg_file_list <- readRDS("RData/Aligned_Data_NoBkg_Smooth.RDS")

PL_NoBkg <- DataNobkg_file_list$PL_NoBkg_Aligned_Smooth
PS_NoBkg <- DataNobkg_file_list$PS_NoBkg_Aligned_Smooth
C2_NoBkg <- DataNobkg_file_list$C2_NoBkg_Aligned_Smooth
C3_NoBkg <- DataNobkg_file_list$C3_NoBkg_Aligned_Smooth
POC_NoBkg <- DataNobkg_file_list$POC_NoBkg_Aligned_Smooth
POE_NoBkg <- DataNobkg_file_list$POE_NoBkg_Aligned_Smooth
JB_NoBkg <- DataNobkg_file_list$JB_NoBkg

AliyahPO_NoBkg <- DataNobkg_file_list$AliyahPO_NoBkg_Smooth

rm(DataNobkg_file_list)

```




# Plotting all samples with "plot" from powdR 

Will plot some spectra to see the samples by using the plot() function from powdR package. The plot.multiXY is designed to provide easy, adaptable plots of multiple XRPD patterns.

```{r plots to share}

ListAllSamples_SHARE_Smooth <- as_multi_xy(list("C2 Sediment Core at 10 cm" = C2_NoBkg$d10cm, "C2 Sediment Core at 15 cm" = C2_NoBkg$d15cm, "C2 Sediment Core at 18 cm" = C2_NoBkg$d18cm, "C2 Sediment Core at 24 cm" = C2_NoBkg$d24cm, "C2 Sediment Core at 28 cm" = C2_NoBkg$d28cm, "C2 Sediment Core at 29 cm" = C2_NoBkg$d29cm, "C2 Sediment Core at 41 cm" = C2_NoBkg$d41cm, "C2 Sediment Core at 44 cm" = C2_NoBkg$d44cm, "C2 Sediment Core at 45 cm" = C2_NoBkg$d45cm, "C2 Sediment Core at 47 cm" = C2_NoBkg$d47cm, "C2 Sediment Core at 49 cm" = C2_NoBkg$d49cm, "C2 Sediment Core at 60 cm" = C2_NoBkg$d60cm, "C2 Sediment Core at 67 cm" = C2_NoBkg$d67cm, "C2 Sediment Core at 70 cm" = C2_NoBkg$d70cm, "C2 Sediment Core at 80 cm" = C2_NoBkg$d80cm, "C2 Sediment Core at 82 cm" = C2_NoBkg$d82cm, "C2 Sediment Core at 84 cm" = C2_NoBkg$d84cm,  "C2 Sediment Core at 85 cm" = C2_NoBkg$d85cm, "C2 Sediment Core at 88 cm" = C2_NoBkg$d88cm, "C2 Sediment Core at 92 cm" = C2_NoBkg$d92cm, "C2 Sediment Core at 96 cm" = C2_NoBkg$d96cm, "C2 Sediment Core at 98 cm" = C2_NoBkg$d98cm, "C2 Sediment Core at 100 cm" = C2_NoBkg$d100cm, "C2 Sediment Core at 102 cm" = C2_NoBkg$d102cm, "C2 Sediment Core at 104 cm" = C2_NoBkg$d104cm, "C2 Sediment Core at 56 cm" = C2_NoBkg$d56cm, "C2 Sediment Core at 58 cm" = C2_NoBkg$d58cm, "C2 Sediment Core at 46 cm" = C2_NoBkg$d46cm, "C2 Sediment Core at 113 cm" = C2_NoBkg$d113cm, "C2 Sediment Core at 115 cm" = C2_NoBkg$d115cm, "C2 Sediment Core at 103 cm" = C2_NoBkg$d103cm, "C2 Sediment Core at 112 cm" = C2_NoBkg$d112cm, "C2 Sediment Core at 74 cm" = C2_NoBkg$d74cm, "C2 Sediment Core at 94 cm" = C2_NoBkg$d94cm, "C3 Sediment Core at 13 cm" = C3_NoBkg$d13cm, "C3 Sediment Core at 18 cm" = C3_NoBkg$d18cm, "C3 Sediment Core at 21 cm" = C3_NoBkg$d21cm, "C3 Sediment Core at 26 cm" = C3_NoBkg$d26cm, "C3 Sediment Core at 30 cm" = C3_NoBkg$d30cm, "C3 Sediment Core at 35 cm" = C3_NoBkg$d35cm, "C3 Sediment Core at 37 cm" = C3_NoBkg$d37cm, "C3 Sediment Core at 52 cm" = C3_NoBkg$d52cm, "C3 Sediment Core at 49 cm" = C3_NoBkg$d49_gs_cm, "C3 Sediment Core at 47 cm" = C3_NoBkg$d47cm, "C3 Sediment Core at 44 cm" = C3_NoBkg$d44cm, "C3 Sediment Core at 39 cm" = C3_NoBkg$d39cm, "Ponce Limestone (rocks close to canal)" = PL_NoBkg$PL_Canal, "Ponce Limestone: PSL C1" = PL_NoBkg$PL_James_LC1, "Ponce Limestone:PSL C2" = PL_NoBkg$PL_James_LC2, "Ponce Limestone: PSL B1" = PL_NoBkg$PL_James_LC2, "Ponce Limestone: PL1" = PL_NoBkg$PL_1_NoHalite, "Ponce Limestone: PL13" = PL_NoBkg$PL_13_NoHalite, "Ponce Limestone: PL15" = PL_NoBkg$PL_15_NoHalite, "Ponce Limestone: PL16" = PL_NoBkg$PL_16_NoHalite, "Ponce Limestone: PL8" = PL_NoBkg$PL_8_NoHalite, "Ponce Limestone: PL5" = PL_NoBkg$PL_5_NoHalite, "Ponce Limestone: PL6" = PL_NoBkg$PL_6_NoHalite, "Ponce Limestone: PL9" = PL_NoBkg$PL_9_NoHalite, "Microbialite from 2022, PST002 Crust)" = PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_002_Sam1_cbb, "Microbialite from 2022, PST002 Middle Black Sediment)" = PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_002_Sam2bsed, "Microbialite from 2022, PST002 Bottom Sediment)" = PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_002_Sam3bottom,  "Microbialite from 2022, PST005 Crust and middle)" = PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_005_Sam1CrustMiddle, "Microbialite from 2022, PST005 Bottom Sediment)" = PS_NoBkg$PS2223_NoBkg_Smooth$PS22T_005_Sam2Bottom, "Microbial Mat from 2022, PS Farm (close to the farm, in the salterns)" = PS_NoBkg$PS2223_NoBkg_Smooth$PS22_Farm, "Microbialite, 2019: upper crust 1" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust1, "Microbialite, 2019: upper crust 2" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust2, "Microbialite, 2019: middle lithified layer" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Middle.Lithifed, "Microbialite, 2019: black layer" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Black.Layer, "Microbialite, 2019: Layer with disordered Ca-dolomite" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_L.below.Black, "Microbialite, 2019: Bottom lithified layer" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Bottom.Lithified, "Microbialite from 2023, black layer" = PS_NoBkg$PS2223_NoBkg_Smooth$PS23_Bottom, "Microbial mat from PO1C, 2019: Crust/green layers" = POC_NoBkg$PO19_01C03_Crust_green, "Microbial mat from PO1C, 2019: Green Layer" = POC_NoBkg$PO19_01C03_Green, "Microbial mat from PO1C, 2019: grains within layers" = POC_NoBkg$PO19_01C03_grains, "Microbial mat from PO1C, 2019: Black layer" = POC_NoBkg$PO19_01C03_Black, "Microbial mat from PO1C, 2019: Bottom sand" = POC_NoBkg$PO19_01C03_Bottom_sand, "Microbial mat from PO1E, 2019: Crust layer" = POE_NoBkg$PO19_01E01_crust.1, "Microbial mat from PO1E, 2019: green layer" = POE_NoBkg$PO19_01E01_green.2, "Microbial mat from PO1E, 2019: brown layer" = POE_NoBkg$PO19_01E01_brown.3, "Microbial mat from PO1E, 2019: black layer" = POE_NoBkg$PO19_01E01_black.4, "Microbial mat from JB, 2019: Crust/green layers" = JB_NoBkg$JB19_Crust.green, "Microbial mat from JB, 2019: pink layer" = JB_NoBkg$JB19_D.pink, "Microbial mat from JB, 2019: Center black layer" = JB_NoBkg$JB19_Center.black, "Sand from PO" = PL_NoBkg$Sand_PO2023, "Sand from Jungle Beach" = PL_NoBkg$Sand_JungleBeach, "Sand close to the canal in Providencia" = PL_NoBkg$Sand_Canal, "Dolomite Standard" = MinStds2797_NoBkg$Dolomite_S2797, "Halite Standard" = MinStds2797_NoBkg$NewHalite_standard_S2797, "Calcite Standard" = MinStds2797_NoBkg$Calcite_S2797, "HMC Standard (from sea urchins)" = MinStds2797_NoBkg$HigMgCal_S2797, "Quartz Standard" = MinStds2797_NoBkg$Quartz_S2797, "Aragonite from Coral" = MinStds2797_NoBkg$Aragonite_coral_standard_S2797, "Gypsum standard" = MinStds2797_NoBkg$Gypsum_S2797))



plot_all <- plot(ListAllSamples_SHARE_Smooth, wavelength = "Cu", xlim = c(10, 58), normalise = T, interactive = T) #xlim = c(34, 36) if you want


plot_all

#Save HTML plot for co-authors

# Convert ggplot to an interactive plotly object
InteractivePlot_all <- ggplotly(plot_all)

# Save as an interactive HTML file
saveWidget(InteractivePlot_all, "plots/InteractivePlot_XRD_all.html", selfcontained = T, libdir = NULL)
#print("Interactive HTML file saved")

```



```{r}

ListCores_Smooth <- as_multi_xy(list("C2 Sediment Core at 10 cm" = C2_NoBkg$d10cm, "C2 Sediment Core at 15 cm" = C2_NoBkg$d15cm, "C2 Sediment Core at 18 cm" = C2_NoBkg$d18cm, "C2 Sediment Core at 24 cm" = C2_NoBkg$d24cm, "C2 Sediment Core at 28 cm" = C2_NoBkg$d28cm, "C2 Sediment Core at 29 cm" = C2_NoBkg$d29cm, "C2 Sediment Core at 41 cm" = C2_NoBkg$d41cm, "C2 Sediment Core at 44 cm" = C2_NoBkg$d44cm, "C2 Sediment Core at 45 cm" = C2_NoBkg$d45cm, "C2 Sediment Core at 47 cm" = C2_NoBkg$d47cm, "C2 Sediment Core at 49 cm" = C2_NoBkg$d49cm, "C2 Sediment Core at 60 cm" = C2_NoBkg$d60cm, "C2 Sediment Core at 67 cm" = C2_NoBkg$d67cm, "C2 Sediment Core at 70 cm" = C2_NoBkg$d70cm, "C2 Sediment Core at 80 cm" = C2_NoBkg$d80cm, "C2 Sediment Core at 82 cm" = C2_NoBkg$d82cm ,"C2 Sediment Core at 84 cm" = C2_NoBkg$d84cm,  "C2 Sediment Core at 85 cm" = C2_NoBkg$d85cm, "C2 Sediment Core at 88 cm" = C2_NoBkg$d88cm, "C2 Sediment Core at 92 cm" = C2_NoBkg$d92cm, "C2 Sediment Core at 96 cm" = C2_NoBkg$d96cm, "C2 Sediment Core at 98 cm" = C2_NoBkg$d98cm, "C2 Sediment Core at 100 cm" = C2_NoBkg$d100cm, "C2 Sediment Core at 102 cm" = C2_NoBkg$d102cm, "C2 Sediment Core at 104 cm" = C2_NoBkg$d104cm, "C2 Sediment Core at 56 cm" = C2_NoBkg$d56cm, "C2 Sediment Core at 58 cm" = C2_NoBkg$d58cm, "C2 Sediment Core at 46 cm" = C2_NoBkg$d46cm, "C2 Sediment Core at 113 cm" = C2_NoBkg$d113cm, "C2 Sediment Core at 115 cm" = C2_NoBkg$d115cm, "C2 Sediment Core at 103 cm" = C2_NoBkg$d103cm, "C2 Sediment Core at 112 cm" = C2_NoBkg$d112cm, "C2 Sediment Core at 74 cm" = C2_NoBkg$d74cm, "C2 Sediment Core at 94 cm" = C2_NoBkg$d94cm, "C3 Sediment Core at 13 cm" = C3_NoBkg$d13cm, "C3 Sediment Core at 18 cm" = C3_NoBkg$d18cm, "C3 Sediment Core at 21 cm" = C3_NoBkg$d21cm, "C3 Sediment Core at 26 cm" = C3_NoBkg$d26cm, "C3 Sediment Core at 30 cm" = C3_NoBkg$d30cm, "C3 Sediment Core at 35 cm" = C3_NoBkg$d35cm, "C3 Sediment Core at 37 cm" = C3_NoBkg$d37cm, "C3 Sediment Core at 52 cm" = C3_NoBkg$d52cm, "C3 Sediment Core at 49 cm" = C3_NoBkg$d49_gs_cm, "C3 Sediment Core at 47 cm" = C3_NoBkg$d47cm, "C3 Sediment Core at 44 cm" = C3_NoBkg$d44cm, "C3 Sediment Core at 39 cm" = C3_NoBkg$d39cm, "Ponce Limestone (rocks close to canal)" = PL_NoBkg$PL_Canal, "Ponce Limestone: PSL C1" = PL_NoBkg$PL_James_LC1, "Ponce Limestone:PSL C2" = PL_NoBkg$PL_James_LC2, "Ponce Limestone: PSL B1" = PL_NoBkg$PL_James_LC2, "Ponce Limestone: PL1" = PL_NoBkg$PL_1_NoHalite, "Ponce Limestone: PL13" = PL_NoBkg$PL_13_NoHalite, "Ponce Limestone: PL15" = PL_NoBkg$PL_15_NoHalite, "Ponce Limestone: PL16" = PL_NoBkg$PL_16_NoHalite, "Ponce Limestone: PL8" = PL_NoBkg$PL_8_NoHalite, "Ponce Limestone: PL5" = PL_NoBkg$PL_5_NoHalite, "Ponce Limestone: PL6" = PL_NoBkg$PL_6_NoHalite, "Ponce Limestone: PL9" = PL_NoBkg$PL_9_NoHalite, "Sand from PO" = PL_NoBkg$Sand_PO2023, "Sand from Jungle Beach" = PL_NoBkg$Sand_JungleBeach, "Sand close to the canal in Providencia" = PL_NoBkg$Sand_Canal, "Dolomite Standard" = MinStds2797_NoBkg$Dolomite_S2797, "Halite Standard" = MinStds2797_NoBkg$NewHalite_standard_S2797, "Calcite Standard" = MinStds2797_NoBkg$Calcite_S2797, "HMC Standard (from sea urchins)" = MinStds2797_NoBkg$HigMgCal_S2797, "Quartz Standard" = MinStds2797_NoBkg$Quartz_S2797, "Aragonite from Coral" = MinStds2797_NoBkg$Aragonite_coral_standard_S2797, "Gypsum standard" = MinStds2797_NoBkg$Gypsum_S2797))


plot_Cores <- plot(ListCores_Smooth, wavelength = "Cu", xlim = c(10, 58), normalise = T, interactive = T) #xlim = c(34, 36) if you want


plot_Cores

#Save HTML plot for co-authors

# Convert ggplot to an interactive plotly object
InteractivePlot_Cores <- ggplotly(plot_Cores)

```



Plotting by samples
```{r Checking plots per sample}

List_JB <- as_multi_xy(list( "JB19_Crust.green" = JB_NoBkg$JB19_Crust.green, "JB19_D.pink" = JB_NoBkg$JB19_D.pink, "JB19_Center.black" = JB_NoBkg$JB19_Center.black, "gypsum" = MinStds2797_NoBkg$Gypsum_S2797, "aragonite" = MinStds2797_NoBkg$Aragonite_coral_standard_S2797, "halite" = MinStds2797_NoBkg$Halite_S2797, "calcite" = MinStds2797_NoBkg$Calcite_S2797, "HMC" = MinStds2797_NoBkg$HigMgCal_S2797, "dolomite" = MinStds2797_NoBkg$Dolomite_S2797))
                             
List_POE <- as_multi_xy(list("PO19_01E01_crust.1" = POE_NoBkg$PO19_01E01_crust.1, "PO19_01E01_green.2" = POE_NoBkg$PO19_01E01_green.2, "PO19_01E01_brown.3" = POE_NoBkg$PO19_01E01_brown.3, "PO19_01E01_black.4" = POE_NoBkg$PO19_01E01_black.4, "aragonite" = MinStds2797_NoBkg$Aragonite_coral_standard_S2797, "halite" = MinStds2797_NoBkg$Halite_S2797, "calcite" = MinStds2797_NoBkg$Calcite_S2797, "HMC" = MinStds2797_NoBkg$HigMgCal_S2797, "dolomite" = MinStds2797_NoBkg$Dolomite_S2797))

List_POC <- as_multi_xy(list("PO19_01C03_Crust_green" = POC_NoBkg$PO19_01C03_Crust_green, "PO19_01C03_Green" = POC_NoBkg$PO19_01C03_Green, "PO19_01C03_grains" = POC_NoBkg$PO19_01C03_grains, "PO19_01C03_Black" = POC_NoBkg$PO19_01C03_Black, "PO19_01C03_Bottom_sand" = POC_NoBkg$PO19_01C03_Bottom_sand, "aragonite" = MinStds2797_NoBkg$Aragonite_coral_standard_S2797, "halite" = MinStds2797_NoBkg$Halite_S2797, "calcite" = MinStds2797_NoBkg$Calcite_S2797, "HMC" = MinStds2797_NoBkg$HigMgCal_S2797, "dolomite" = MinStds2797_NoBkg$Dolomite_S2797))

List_PS19 <- as_multi_xy(list("PS19_Upper.crust1" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust1, "PS19_Upper.crust2" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust2, "PS19_Middle.Lithifed" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Middle.Lithifed, "PS19_Black.Layer" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Black.Layer, "PS19_L.below.Black" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_L.below.Black, "PS19_Bottom.Lithified" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Bottom.Lithified, "aragonite" = MinStds2797_NoBkg$Aragonite_coral_standard_S2797, "halite" = MinStds2797_NoBkg$Halite_S2797, "calcite" = MinStds2797_NoBkg$Calcite_S2797, "HMC" = MinStds2797_NoBkg$HigMgCal_S2797, "dolomite" = MinStds2797_NoBkg$Dolomite_S2797))

plot(List_PS19, wavelength = "Cu", normalise = T, interactive = T) #Choose whichever you want to plot
```



# Plotting Individual

```{r}

#ForPaper
ForPaper <- plot(PS_NoBkg$PS19_NoBkg_Smooth, xlim = c(28, 32), wavelength = "Cu", normalise = T) +
  theme_bw()

plot(PL_NoBkg$Sand_Canal, wavelength = "Cu", normalise = T) +
  theme_bw()

#To viz
#plot(as_multi_xy(c(PS_NoBkg, PL_NoBkg, C3_NoBkg, MinStds2797_NoBkg)), wavelength = "Cu", xlim = c(20, 55), normalise = T, interactive = T)

# xlim = c(28, 32), For Zoom regions.


# Save as SVG
ggsave("plots/PS_ZoomedRegion.svg", plot = ForPaper, width = 10, height = 8)
```



# Plotting with ggplot

In the following markdown, I will create a code that will plot only one set of samples. The only thing you will havr to change is the plot_data dataframe information. 

```{r ggplot for manuscripts}


plot_data <- as_multi_xy(PL_NoBkg) #Change the sample list to whichever one you desire. 


# Pivot the data from wide to long format for plotting
df_long <- pivot_longer(PL_NoBkg, 
                        cols = -tth, 
                        names_to = "sample", 
                        values_to = "counts")

# Assuming df_long has columns: tth, counts, and sample
df_norm <- df_long %>% 
  group_by(sample) %>% 
  mutate(norm_counts = counts / max(counts, na.rm = TRUE)) %>% 
  ungroup()

plotIllustrator <- ggplot(df_norm, aes(x = tth, y = norm_counts, color = sample)) +
  geom_line(size = 0.2) +
  labs(title = "XRD Spectra", 
       x = "2θ", 
       y = "Intensity", 
       color = "Sample") +
  scale_x_continuous(limits = c(10, 58), 
                     breaks = seq(10, 58, by = 5), 
                     minor_breaks = seq(10, 58, by = 1),
                     guide = guide_axis_minor()) +
  scale_y_continuous(limits = c(min(df_norm$norm_counts), max(df_norm$norm_counts))) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),  
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.background = element_rect(fill = "white", color = NA),
    axis.ticks = element_line(color = "black"),
    axis.ticks.length = unit(0.2, "cm"),
    axis.text.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5)  # Center the title
  ) +
  geom_vline(xintercept = 22, linetype = "dashed", color = "black", size = 0.2) +
  annotate("text", x = 22, y = max(df_norm$norm_counts, na.rm = TRUE) * 0.95,
           label = "d-101", angle = 90, vjust = -0.6) +
  geom_vline(xintercept = 31, linetype = "dashed", color = "black", size = 0.2) +
  annotate("text", x = 31, y = max(df_norm$norm_counts, na.rm = TRUE) * 0.95,
           label = "d-104", angle = 90, vjust = -0.6) +
  geom_vline(xintercept = 35.3, linetype = "dashed", color = "black", size = 0.2) +
  annotate("text", x = 35.3, y = max(df_norm$norm_counts, na.rm = TRUE) * 0.95,
           label = "d-015", angle = 90, vjust = -0.6) +
geom_vline(xintercept = 43.8, linetype = "dashed", color = "black", size = 0.2) +
  annotate("text", x = 43.8, y = max(df_norm$norm_counts, na.rm = TRUE) * 0.95,
           label = "d-021", angle = 90, vjust = -0.6)



# Convert to an interactive plot
Interactive <- ggplotly(plotIllustrator)

plotIllustrator
```




```{r Plotting only specific regions}

plotIllustrator_Specific <- ggplot(PL_NoBkg, aes(x = tth, y = norm_counts, color = sample)) +
  geom_line(size = 0.2) +
  labs(title = "XRD Spectra", 
       x = "2θ", 
       y = "Intensity", 
       color = "Sample") +
  scale_x_continuous(limits = c(20, 24), 
                     breaks = seq(20, 24, by = 1), 
                     minor_breaks = seq(20, 24, by = 0.5),
                     guide = guide_axis_minor()) +
  scale_y_continuous(limits = c(0, max(df_norm$norm_counts))) +
  theme_minimal() +
  theme(
    panel.grid.major = element_blank(),  
    panel.grid.minor = element_blank(),  
    panel.border = element_rect(color = "black", fill = NA, linewidth = 1),
    panel.background = element_rect(fill = "white", color = NA),
    axis.ticks = element_line(color = "black"),
    axis.ticks.length = unit(0.2, "cm"),
    axis.text.x = element_text(margin = margin(t = 10)),
    plot.title = element_text(hjust = 0.5))

# Convert to an interactive plot
Interactive <- ggplotly(plotIllustrator_Specific)

plotIllustrator_Specific
    
    
```




# Saving Plots for Illustator and as interactive

```{r Saving Plots}

# Convert ggplot to an interactive plotly object
InteractivePlot <- ggplotly(plotIllustrator)

# Save as an interactive HTML file
saveWidget(InteractivePlot, "plots/InteractivePlot_stds.html", selfcontained = T, libdir = NULL)
print("Interactive HTML file saved")


# Save as SVG (for Illustrator editing) #Change the name everytime you change the dataset
ggsave(filename = "plots/plotIllustrator_SciRep.svg", plot = plot_all, device = "svg", width = 13, height = 13, dpi = 300)
print("SVG file saved")


# Save as TIFF (high-resolution for publication)
ggsave(filename = "plots/plotIllustrator_stds.tiff", plot = plot_all, device = "tiff", width = 13, height = 13, dpi = 600, compression = "lzw")
print("TIFF file saved")


```




