---
title: 'XRD Code 3: MgCa % calculation'
author: "Bryan Rodriguez"
date: "2025-03-23"
output: html_document
---

In the following section, I will implement some calculations provided by Luis Gonzalez, from Arvidson & Mackenzie, 1999 and Lumsden and Chimahusky, 1980 to calculate the mol % of MgCO3 in the calcites and dolomites of the samples here to see if they are high Mg-calcites or not. 

What I will do is that I will create a separate function to calculate the Mg in calcites and Mg in dolomites. There are several types of samples included in my data: some of them have only dolomite, other have only calcite, others have a mix and other have all those three but not crystalline. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
setwd("~/OneDrive - University of Kansas/Bryan-MS_PhD/PHD-PROJECT/PuertoRico/CHAPTER3/XRD/R_plotting")

```


```{r load packages, include=F, ECHO =F}
library(powdR)
library(ggplot2)
library(ggsci)
library(RColorBrewer)
library(plotly)
library(dplyr)
library(tibble)
library(purrr)
library(ggpubr)
library(lubridate)
library(patchwork)
library(baseline)
library(EEM)
library(data.table)
library(tidyverse)
library(conflicted)
library(stringr)
```



# Functions

The following functions will be used to get the d-spacing of crystalline phases by getting that maxium value from the highest peak
```{r Mol% calculation Function for crystalline calcites}
calculate_calcite_for_list <- function(df_listCal, calcite_range) {
  results_listCal <- list()
  
  for (i in seq_along(df_listCal)) {
    sample_name <- names(df_listCal)[i]
    sample_df <- df_listCal[[i]]
    
    # Select rows where "2theta" is in the specified range for Calcite
    selected_dataCal <- sample_df[sample_df$tth >= calcite_range[1] & sample_df$tth <= calcite_range[2], ]
    
    if (nrow(selected_dataCal) == 0) {
      # If no data in the specified range, return an empty data frame
      next
    }
    
    # Find the index of the row with the highest counts
    max_count_index <- which.max(selected_dataCal$counts)
    
    # Get the tth value with the highest counts
    max_tth <- selected_dataCal$tth[max_count_index]
    
    # Calculate DivSum and dSpacing for Calcite using the max tth
    DivSumCal <- sum((selected_dataCal$counts * max_tth)) / sum(selected_dataCal$counts)
    dSpacingCal <- (1.54056) / (2 * sin((pi / 180) * (DivSumCal / 2)))
    
    # Calculate values using equations for Calcite
    equation1 <- (-363.96 * dSpacingCal) + 1104.5 # Arvidson & Mackenzie, 1999
    equation2 <- 100 - ((333.33 * dSpacingCal) - 911.99) # Lumsden and Chimahusky, 1980
    
    # Determine the Calcite Type based on the calculated values
    calcite_type1 <- ifelse(equation1 >= 0 & equation1 < 4, "Low Mg-Calcite",
                            ifelse(equation1 >= 4 & equation1 < 12, "Intermediate Mg-Calcite",
                                   ifelse(equation1 >= 12 & equation1 <= 30, "High Mg-Calcite", 
                                          ifelse(equation1 >= 31 & equation1 <= 44, "Very High Mg-Calcite", NA))))
    
    calcite_type2 <- ifelse(equation2 >= 0 & equation2 < 4, "Low Mg-Calcite",
                            ifelse(equation2 >= 4 & equation2 < 12, "Intermediate Mg-Calcite",
                                   ifelse(equation2 >= 12 & equation2 <= 30, "High Mg-Calcite", 
                                  ifelse(equation2 >= 31 & equation2 <= 44, " Very High Mg-Calcite", NA))))
    
    # Calculate mol_percentage_CaCO3
    mol_percentage_CaCO3_1 <- 100 - equation1
    mol_percentage_CaCO3_2 <- 100 - equation2
    
    # Create data frames for the results for Calcite based on both equations
    result_dfCal1 <- data.frame(
      "Sample" = sample_name,
      "Mineral_Type" = "Calcite",
      "mol_perc_MgCo3_calc_eq" = "Arvidson & Mackenzie, 1999",
      "mol_percentage_MgCO3" = round(equation1, 2),
      "mol_percentage_CaCO3" = round(mol_percentage_CaCO3_1, 2),
      "Mineral_Phase" = calcite_type1,
      "Max_tth_in_range" = max_tth,
      "dSpacing" = dSpacingCal
    )
    
    result_dfCal2 <- data.frame(
      "Sample" = sample_name,
      "Mineral_Type" = "Calcite",
      "mol_perc_MgCo3_calc_eq" = "Lumsden and Chimahusky, 1980",
      "mol_percentage_MgCO3" = round(equation2, 2),
      "mol_percentage_CaCO3" = round(mol_percentage_CaCO3_2, 2),
      "Mineral_Phase" = calcite_type2,
      "Max_tth_in_range" = max_tth,
      "dSpacing" = dSpacingCal
    )
    
    results_listCal[[i]] <- rbind(result_dfCal1, result_dfCal2) # Store both results for the sample
  }
  
  final_result_ListCal <- do.call(rbind, results_listCal)
  row.names(final_result_ListCal) <- NULL
  
  return(final_result_ListCal)
}
```


```{r Mol% calculation Function for crystalline dolomites}

calculate_dolomite_for_list <- function(df_listDol, dolomite_range) {
  results_listDol <- list()
  
  for (i in seq_along(df_listDol)) {
    sample_name <- names(df_listDol)[i]
    sample_df <- df_listDol[[i]]
    
    # Select rows where "2theta" is in the specified range for Dolomite
    selected_dataDol <- sample_df[sample_df$tth >= dolomite_range[1] & sample_df$tth <= dolomite_range[2], ]
    
    if (nrow(selected_dataDol) == 0) {
      # If no data in the specified range, return an empty data frame
      next
    }
    
    # Find the index of the row with the highest counts
    max_count_index <- which.max(selected_dataDol$counts)
    
    # Get the tth value with the highest counts
    max_tth <- selected_dataDol$tth[max_count_index]
    
    # Calculate DivSum and dSpacing for Dolomite using the max tth
    DivSumDol <- sum((selected_dataDol$counts * max_tth)) / sum(selected_dataDol$counts)
    dSpacingDol <- (1.54056) / (2 * sin((pi / 180) * (DivSumDol / 2)))
    
    # Calculate values using equations for Dolomite
    equation1 <- (-363.96 * dSpacingDol) + 1104.5 # Arvidson & Mackenzie, 1999
    equation2 <- 100 - ((333.33 * dSpacingDol) - 911.99) # Lumsden and Chimahusky, 1980
    
    # Determine the Dolomite Type based on the calculated values
    dolomite_type1 <- ifelse(equation1 >= 45 & equation1 < 49, "Ca-dolomite",
                            ifelse(equation1 >= 49.4 & equation1 <= 50.4, "Stoichiometric Dolomite",
                                   ifelse(equation1 >= 50.5, "Mg-Dolomite", "Very High Mg-Calcite")))
    
    dolomite_type2 <- ifelse(equation2 >= 45 & equation2 < 49, "Ca-dolomite",
                            ifelse(equation2 >= 49.4 & equation2 <= 50.4, "Stoichiometric Dolomite",
                                   ifelse(equation2 >= 50.5, "Mg-Dolomite", "Very High Mg-Calcite")))
    
    # Calculate mol_percentage_CaCO3
    mol_percentage_CaCO3_1 <- 100 - equation1
    mol_percentage_CaCO3_2 <- 100 - equation2
    
    # Create data frames for the results for Dolomite based on both equations
    result_dfDol1 <- data.frame(
      "Sample" = sample_name,
      "Mineral_Type" = "Dolomite",
      "mol_perc_MgCo3_calc_eq" = "Arvidson & Mackenzie, 1999",
      "mol_percentage_MgCO3" = round(equation1, 2),
      "mol_percentage_CaCO3" = round(mol_percentage_CaCO3_1, 2),
      "Mineral_Phase" = dolomite_type1,
      "Max_tth_in_range" = max_tth,
      "dSpacing" = dSpacingDol
    )
    
    result_dfDol2 <- data.frame(
      "Sample" = sample_name,
      "Mineral_Type" = "Dolomite",
      "mol_perc_MgCo3_calc_eq" = "Lumsden and Chimahusky, 1980",
      "mol_percentage_MgCO3" = round(equation2, 2),
      "mol_percentage_CaCO3" = round(mol_percentage_CaCO3_2, 2),
      "Mineral_Phase" = dolomite_type2,
      "Max_tth_in_range" = max_tth,
      "dSpacing" = dSpacingDol
    )
    
    results_listDol[[i]] <- rbind(result_dfDol1, result_dfDol2) # Store both results for the sample
  }
  
  final_result_ListDol <- do.call(rbind, results_listDol)
  row.names(final_result_ListDol) <- NULL
  
  return(final_result_ListDol)
}

```

```{r Mol% calculation Function for NON-crystalline calcites}
calculate_calcite_for_list_NonCrystalline <- function(df_listCal, calcite_range) {
  results_listCal <- list()
  
  for (i in seq_along(df_listCal)) {
    sample_name <- names(df_listCal)[i]
    sample_df <- df_listCal[[i]]
    
    # Select rows where "2theta" is in the specified range for Calcite
    selected_dataCal <- sample_df[sample_df$tth >= calcite_range[1] & sample_df$tth <= calcite_range[2], ]
    
    if (nrow(selected_dataCal) == 0) {
      # If no data in the specified range, return an empty data frame
      next
    }
    
    
    # Find the index of the row with the highest counts
    max_count_index <- which.max(selected_dataCal$counts)
    
    # Get the tth value with the highest counts
    max_tth <- selected_dataCal$tth[max_count_index]
    
    # Calculate DivSum and dSpacing for Calcite using the max tth
    DivSumCal <- sum((selected_dataCal$counts * selected_dataCal$tth)) / sum(selected_dataCal$counts)
    dSpacingCal <- (1.54056) / (2 * sin((pi / 180) * (DivSumCal / 2)))
    
    # Calculate values using equations for Calcite
    equation1 <- (-363.96 * dSpacingCal) + 1104.5 # Arvidson & Mackenzie, 1999
    equation2 <- 100 - ((333.33 * dSpacingCal) - 911.99) # Lumsden and Chimahusky, 1980
    
    # Determine the Calcite Type based on the calculated values
    calcite_type1 <- ifelse(equation1 >= 0 & equation1 < 4, "Low Mg-Calcite",
                            ifelse(equation1 >= 4 & equation1 < 12, "Intermediate Mg-Calcite",
                                   ifelse(equation1 >= 12 & equation1 <= 30, "High Mg-Calcite", 
                                          ifelse(equation1 >= 31 & equation1 <= 44, "Very High Mg-Calcite", NA))))
    
    calcite_type2 <- ifelse(equation2 >= 0 & equation2 < 4, "Low Mg-Calcite",
                            ifelse(equation2 >= 4 & equation2 < 12, "Intermediate Mg-Calcite",
                                   ifelse(equation2 >= 12 & equation2 <= 30, "High Mg-Calcite", 
                                  ifelse(equation2 >= 31 & equation2 <= 44, " Very High Mg-Calcite", NA))))
    
    # Calculate mol_percentage_CaCO3
    mol_percentage_CaCO3_1 <- 100 - equation1
    mol_percentage_CaCO3_2 <- 100 - equation2
    
    # Create data frames for the results for Calcite based on both equations
    result_dfCal1 <- data.frame(
      "Sample" = sample_name,
      "Mineral_Type" = "Calcite",
      "mol_perc_MgCo3_calc_eq" = "Arvidson & Mackenzie, 1999",
      "mol_percentage_MgCO3" = round(equation1, 2),
      "mol_percentage_CaCO3" = round(mol_percentage_CaCO3_1, 2),
      "Mineral_Phase" = calcite_type1,
      "Calculated_dSpacing" = dSpacingCal,
      "Max_2theta_in_range" = max_tth
    )
    
    result_dfCal2 <- data.frame(
      "Sample" = sample_name,
      "Mineral_Type" = "Calcite",
      "mol_perc_MgCo3_calc_eq" = "Lumsden and Chimahusky, 1980",
      "mol_percentage_MgCO3" = round(equation2, 2),
      "mol_percentage_CaCO3" = round(mol_percentage_CaCO3_2, 2),
      "Mineral_Phase" = calcite_type2,
      "Calculated_dSpacing" = dSpacingCal,
      "Max_2theta_in_range" = max_tth
    )
    
    results_listCal[[i]] <- rbind(result_dfCal1, result_dfCal2) # Store both results for the sample
  }
  
  final_result_ListCal <- do.call(rbind, results_listCal)
  row.names(final_result_ListCal) <- NULL
  
  return(final_result_ListCal)
}
```

```{r Mol% calculation Function for NON-crystalline dolomites}

calculate_dolomite_for_list_NonCrystalline <- function(df_listDol, dolomite_range) {
  results_listDol <- list()
  
  for (i in seq_along(df_listDol)) {
    sample_name <- names(df_listDol)[i]
    sample_df <- df_listDol[[i]]
    
    # Select rows where "2theta" is in the specified range for Dolomite
    selected_dataDol <- sample_df[sample_df$tth >= dolomite_range[1] & sample_df$tth <= dolomite_range[2], ]
    
    if (nrow(selected_dataDol) == 0) {
      # If no data in the specified range, return an empty data frame
      next
    }
    
    
     # Find the index of the row with the highest counts
    max_count_index <- which.max(selected_dataDol$counts)
    
    # Get the tth value with the highest counts
    max_tth <- selected_dataDol$tth[max_count_index]
    
    # Calculate DivSum and dSpacing for Dolomite using the max tth
    DivSumDol <- sum((selected_dataDol$counts * selected_dataDol$tth)) / sum(selected_dataDol$counts)
    dSpacingDol <- (1.54056) / (2 * sin((pi / 180) * (DivSumDol / 2)))
    
    # Calculate values using equations for Dolomite
    equation1 <- (-363.96 * dSpacingDol) + 1104.5 # Arvidson & Mackenzie, 1999
    equation2 <- 100 - ((333.33 * dSpacingDol) - 911.99) # Lumsden and Chimahusky, 1980
    
    # Determine the Dolomite Type based on the calculated values
    dolomite_type1 <- ifelse(equation1 >= 45 & equation1 < 49, "Ca-dolomite",
                            ifelse(equation1 >= 49.4 & equation1 <= 50.4, "Stoichiometric Dolomite",
                                   ifelse(equation1 >= 50.5, "Mg-Dolomite", "Very High Mg-Calcite")))
    
    dolomite_type2 <- ifelse(equation2 >= 45 & equation2 < 49, "Ca-dolomite",
                            ifelse(equation2 >= 49.4 & equation2 <= 50.4, "Stoichiometric Dolomite",
                                   ifelse(equation2 >= 50.5, "Mg-Dolomite", "Very High Mg-Calcite")))
    
    # Calculate mol_percentage_CaCO3
    mol_percentage_CaCO3_1 <- 100 - equation1
    mol_percentage_CaCO3_2 <- 100 - equation2
    
    # Create data frames for the results for Dolomite based on both equations
    result_dfDol1 <- data.frame(
      "Sample" = sample_name,
      "Mineral_Type" = "Dolomite",
      "mol_perc_MgCo3_calc_eq" = "Arvidson & Mackenzie, 1999",
      "mol_percentage_MgCO3" = round(equation1, 2),
      "mol_percentage_CaCO3" = round(mol_percentage_CaCO3_1, 2),
      "Mineral_Phase" = dolomite_type1,
      "Calculated_dSpacing" = dSpacingDol,
      "Max_2theta_in_range" = max_tth
    )
    
    result_dfDol2 <- data.frame(
      "Sample" = sample_name,
      "Mineral_Type" = "Dolomite",
      "mol_perc_MgCo3_calc_eq" = "Lumsden and Chimahusky, 1980",
      "mol_percentage_MgCO3" = round(equation2, 2),
      "mol_percentage_CaCO3" = round(mol_percentage_CaCO3_2, 2),
      "Mineral_Phase" = dolomite_type2,
      "Calculated_dSpacing" = dSpacingDol,
      "Max_2theta_in_range" = max_tth
    )
    
    results_listDol[[i]] <- rbind(result_dfDol1, result_dfDol2) # Store both results for the sample
  }
  
  final_result_ListDol <- do.call(rbind, results_listDol)
  row.names(final_result_ListDol) <- NULL
  
  return(final_result_ListDol)
}


```



```{r To reset data without erasing functions}
rm(list = ls()[!sapply(ls(), function(x) is.function(get(x)))])
```


# Data

```{r Load Data}


minstds_file_list <- readRDS("RData/min_standards.RDS")
MinStds8389_NoBkg <-minstds_file_list$MinStds8389_NoBkg
MinStds2797_NoBkg <- minstds_file_list$MinStds2797_NoBkg
rm(minstds_file_list)

DataNobkg_file_list <- readRDS("RData/Aligned_Data_NoBkg_Smooth.RDS")

PL_NoBkg <- DataNobkg_file_list$PL_NoBkg
PS_NoBkg <- DataNobkg_file_list$PS_NoBkg
C2_NoBkg <- DataNobkg_file_list$C2_NoBkg
C3_NoBkg <- DataNobkg_file_list$C3_NoBkg
POC_NoBkg <- DataNobkg_file_list$POC_NoBkg
POE_NoBkg <- DataNobkg_file_list$POE_NoBkg
JB_NoBkg <- DataNobkg_file_list$JB_NoBkg

rm(DataNobkg_file_list)

```

##Prepare data for functions


```{r Function inputs}

# Below here are the specified tth ranges for Calcite and Dolomite to be used in the in the crystalline fucntions
calcite_tth_range <- c(28.5, 30.4)
dolomite_tth_range <- c(30.6, 31.1)

#For non crystalline functions
LowMgCalcite_tth_range <- c(29.1, 29.6)
Mgcalcite_tth_range_NonCrys <- c(29.77, 30.14)
dolomite_tth_range_NonCrys <- c(30.6, 31.33)

##For PS-19, it will be: 
MicroLowMgCalcite <- c(29.44, 29.65)
MicroHighMgCalcite <- c(29.85,29.97)
MicroVHMC <- c(30.3, 30.9)


##For POE1
PO_Mgcal <- c(29.44, 30.70)

##For JB
JB_Mgcal <- c(29.80, 30.62)

##PS19Crust

PS19Crust_Mgcal <- c(29.1, 29.4)

```


```{r Ponce Limestone Data With Standards}

#Crystalline calcite and dolomite phases
#Let's start with the crystalline samples that contained calcite. Those are Pl 1, 13, 9 and 16:
MolMgList_PLCalCrystalline <- as_multi_xy(list("Pl-1" = PL_NoBkg$PL_1_NoHalite, "Pl-16" = PL_NoBkg$PL_16_NoHalite, "Pl-13" = PL_NoBkg$PL_13_NoHalite, "Pl-9" = PL_NoBkg$PL_9_NoHalite, "Calcite Standard" = MinStds2797_NoBkg$Calcite_S2797, "High Mg Calcite Standard" = MinStds2797_NoBkg$HigMgCal_S2797))

#Now lets do a list for those with dolomite. Those are 15, 16 (again), 5, 6, 8, canal, B1, C1 and C2:
MolMgList_PLDolCrystalline <- as_multi_xy(list("Pl-15" = PL_NoBkg$PL_15_NoHalite, "Pl-16" = PL_NoBkg$PL_16_NoHalite, "Pl-5" = PL_NoBkg$PL_5_NoHalite, "Pl-8" = PL_NoBkg$PL_8_NoHalite, "Pl-C2" = PL_NoBkg$PL_James_LC2, "Pl-Canal" = PL_NoBkg$PL_Canal, "Pl-B1" = PL_NoBkg$PL_James_LB1, "Pl-C1" = PL_NoBkg$PL_James_LC1, "Dolomite Standard" = MinStds2797_NoBkg$Dolomite_S2797))

#Non crystalline phases

MolMgList_PLCalNonCrystalline <- as_multi_xy(list("Pl-15" = PL_NoBkg$PL_15_NoHalite, "Pl-5" = PL_NoBkg$PL_5_NoHalite, "Pl-6" = PL_NoBkg$PL_6_NoHalite, "Pl-Canal" = PL_NoBkg$PL_Canal))

MolMgList_PLDolNonCrystalline <- as_multi_xy(list("Pl-6" = PL_NoBkg$PL_6_NoHalite, "Dolomite Standard" = MinStds2797_NoBkg$Dolomite_S2797)) #Only PL-6 has a different dolomite peak than others.

```


```{r Sediment Core data with Standards}

#C2
#Let's start with the crystalline samples that contained calcite.
MolMgList_C2CalCrystalline <- as_multi_xy(list("C2 112 cm" = C2_NoBkg$d112cm, "C2 103 cm" = C2_NoBkg$d103cm, "Calcite Standard" = MinStds2797_NoBkg$Calcite_S2797, "High Mg Calcite Standard" = MinStds2797_NoBkg$HigMgCal_S2797))

#C3
MolMgList_C3CalCrystalline <- as_multi_xy(list("C3 39 cm" = C3_NoBkg$d39cm, "C3 42 cm" = C3_NoBkg$d42cm, "C3 44 cm" = C3_NoBkg$d44cm, "C3 47 cm" = C3_NoBkg$d47cm, "C3 49 cm" = C3_NoBkg$d49_gs_cm, "C3 52 cm" = C3_NoBkg$d52cm,"Calcite Standard" = MinStds2797_NoBkg$Calcite_S2797, "High Mg Calcite Standard" = MinStds2797_NoBkg$HigMgCal_S2797))

```


```{r Beach Sand data with Standards}

MolMgList_SandsCalNonCrystalline <- as_multi_xy(list("Sand - Canal" = PL_NoBkg$Sand_Canal, "Sand - Jungle Beach" = PL_NoBkg$Sand_JungleBeach, "Sand - PO" = PL_NoBkg$Sand_PO2023, "Calcite Standard" = MinStds2797_NoBkg$Calcite_S2797))

MolMgList_SandsMgCalNonCrystalline <- as_multi_xy(list("Sand - Canal" = PL_NoBkg$Sand_Canal, "Sand - Jungle Beach" = PL_NoBkg$Sand_JungleBeach, "Sand - PO" = PL_NoBkg$Sand_PO2023, "High Mg Calcite Standard" = MinStds2797_NoBkg$HigMgCal_S2797))

MolMgList_SandsDolNonCrystalline <- as_multi_xy(list("Sand - PO" = PL_NoBkg$PL_6_NoHalite, "Dolomite Standard" = MinStds2797_NoBkg$Dolomite_S2797)) 

```


```{r Microbial deposits with standards}

### POE1
MolMgCaList_POE_CalNonCrystalline <-  as_multi_xy(list("PO-01E crust (2019)" = POE_NoBkg$PO19_01E01_green.2, "PO-01E black layer (2019)" = POE_NoBkg$PO19_01E01_black.4, "calcite Standard" = MinStds2797_NoBkg$Calcite_S2797, "HMC Standard" = MinStds2797_NoBkg$HigMgCal_S2797))

## JB-NW
MolMgCaList_JB_CalNonCrystalline <- as_multi_xy(list("JB-NW Black Layer" = JB_NoBkg$JB19_Center.black, "calcite Standard" = MinStds2797_NoBkg$Calcite_S2797, "HMC Standard" = MinStds2797_NoBkg$HigMgCal_S2797))


## PS19
MolMgCaList_PS19Crust_CalNonCrystalline <- as_multi_xy(list("PS19 Upper Crust 1" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust2, "PS19 Upper Crust 2" = PS_NoBkg$PS19_NoBkg_Smooth$PS19_Upper.crust1, "calcite Standard" = MinStds2797_NoBkg$Calcite_S2797, "HMC Standard" = MinStds2797_NoBkg$HigMgCal_S2797))



```



#Calculate Mol % of Mg and Ca in samples

##Ponce Limestone

```{r Calculation}

#CRYSTALLINE
#Calculate % in Calcites
resultsCalPL_Crys <- calculate_calcite_for_list(MolMgList_PLCalCrystalline, calcite_tth_range)

#Calculate % in Dolomites
resultsDolPL_Crys <- calculate_dolomite_for_list(MolMgList_PLDolCrystalline, dolomite_tth_range)

#NONCRYSTALLINE

# Combine results
resultsCalPL_NoCrys <- calculate_calcite_for_list_NonCrystalline(MolMgList_PLCalNonCrystalline, calcite_tth_range_NonCrys)
#resultsSecondCalPL_NoCrys <- calculate_calcite_for_list_NonCrystalline(MolMgList_PLCalNonCrystalline, SecondCalcite_tth_range)
resultsDolPL_NoCrys <- calculate_dolomite_for_list_NonCrystalline(MolMgList_PLDolNonCrystalline, dolomite_tth_range_NonCrys)

PL_FinalList_NoCrys <- rbind(resultsCalPL_NoCrys, resultsDolPL_NoCrys) #include resultsSecondCalPL_NoCrys[c(7,8),] if you are going to use it. 

```

```{r Combine and Save}

# Combine results
PL_FinalList <- rbind(resultsCalPL_Crys, resultsDolPL_Crys)

# Export the final results to a CSV file
write.csv(PL_FinalList, "OutputData/files/PonceLimestone_MgCa_molperc.csv", row.names = FALSE)
cat("Results have been saved to 'PonceLimestone_MgCa_molperc_ForJames.csv'.\n")

```


##Sediment Cores

```{r Calculation}

##Calculate for C2 (only calcite)
resultsCalC2_Crys <- calculate_calcite_for_list(MolMgList_C2CalCrystalline, calcite_tth_range)

##C3 (both calcites and dolomites)
resultsCalC3_Crys <- calculate_calcite_for_list(MolMgList_C3CalCrystalline, calcite_tth_range)
resultsDolC3_Crys <- calculate_dolomite_for_list(MolMgList_C3CalCrystalline, dolomite_tth_range)

```

```{r Combine Results and Save}

# Combine results
Cores_FinalList <- rbind(resultsCalC2_Crys, resultsCalC3_Crys, resultsDolC3_Crys)

# Export the final results to a CSV file
write.csv(Cores_FinalList, "SedimentCores_MgCa_molperc.csv", row.names = FALSE)
cat("Results have been saved to 'SedimentCores_MgCa_molperc.csv'.\n")

```


##Beach Sand

```{r Calculation}

# Combine results
resultsCalSands_NoCrys <- calculate_calcite_for_list(MolMgList_SandsCalNonCrystalline, LowMgCalcite_tth_range)
resultsMgCalSands_NoCrys <- calculate_calcite_for_list(MolMgList_SandsMgCalNonCrystalline, Mgcalcite_tth_range_NonCrys)
resultsDolSands_NoCrys <- calculate_dolomite_for_list(MolMgList_SandsDolNonCrystalline, dolomite_tth_range_NonCrys)

```

```{r Combined and save}

Sands_FinalList_NoCrys <- rbind(resultsCalSands_NoCrys, resultsMgCalSands_NoCrys, resultsDolSands_NoCrys)

#Save
write.csv(Sands_FinalList_NoCrys, "BeachSand_MgCa_molperc.csv", row.names = FALSE)

```



##Microbialites

```{r Calculation}

## POE1
HighMgresultsCal_POE_NoCrys <- calculate_calcite_for_list(MolMgCaList_POE_CalNonCrystalline, PO_Mgcal)

## JB
HighMgresultsCal_JB_NoCrys <- calculate_calcite_for_list(MolMgCaList_JB_CalNonCrystalline, JB_Mgcal)

## PS19 crusts

HighMgresultsCal_PS19Crust_NoCrys <- calculate_calcite_for_list_NonCrystalline(MolMgCaList_PS19Crust_CalNonCrystalline, PS19Crust_Mgcal)



###PS19 balck layer
HighMgresultsCalMicro_NoCrys <- calculate_calcite_for_list_NonCrystalline(MolMgList_MicroCalNonCrystalline, MicroHighMgCalcite)

```


```{r Combined and save}

POJB_FinalList_NoCrys <- rbind(HighMgresultsCal_POE_NoCrys, HighMgresultsCal_JB_NoCrys)

print(POJB_FinalList_NoCrys)
write.csv(HighMgresultsCal_PS19Crust_NoCrys, "OutputData/files/PS19Crust_MgCa_content.csv", row.names = FALSE)


```



