---
title: 'Code 2: isoOrg'
author: "Bryan Rodriguez-Colon"
date: "2025-03-24"
output: html_document
---
***Disclaimer: This code was designed by Bryan Rodriguez-Colon with the help of chatGPT-5.***


# Overview

This script processes and visualizes organic geochemical data (δ¹³C<sub>org</sub> and TOC/TN ratios) from sediment cores and microbialite samples collected in the Guánica coastal lagoon system, southwestern Puerto Rico.

The script generates both static and interactive crossplots used in the associated *Sedimentology* manuscript to interpret organic matter sources and paleoenvironmental conditions.

In addition, statistical comparisons are performed to evaluate the geochemical distinctiveness of overwash deposits.



## Associated manuscript

Rodríguez-Colón, B. J., et al. (2026)  
*Paleoenvironmental and Regional Controls on Microbialite Development and Distribution in a Coastal Lagoon of Southwestern Puerto Rico*  
Sedimentology


## Input data

The script reads organic geochemical data from: `InputData/Org_isotope_Guanica_data.csv`


The dataset includes:

- δ¹³C<sub>org</sub> (‰ VPDB)  
- Total organic carbon (TOC)  
- Total nitrogen (TN)  
- TOC/TN ratios  
- Sample depth and stratigraphic identifiers  

Rows with missing nitrogen or δ¹⁵N values are removed to ensure consistency in geochemical interpretation.

---

## Statistical analysis

A two-sample t-test is performed to evaluate whether the overwash deposit layer (ZC3C) is geochemically distinct from other stratigraphic intervals.

Comparisons include:

- TOC/TN ratios  
- δ¹³C<sub>org</sub> values  

This analysis provides a quantitative assessment of whether the overwash event represents a statistically distinct organic matter source or depositional regime.

---

## Plot structure

The primary plot displays:

**X-axis**  
TOC/TN ratio

**Y-axis**  
δ¹³C<sub>org</sub> (‰ VPDB)

This crossplot is widely used to infer organic matter sources, including:

- terrestrial C₃ plants  
- terrestrial C₄ plants  
- marine organic matter  
- microbial biomass  

(Lamb et al., 2006).

---

## Environmental interpretation fields

### Rectangular regions (Lamb et al., 2006)

Rectangular fields are defined to represent generalized organic matter sources:

- C₃ plants  
- C₄ plants  
- marine particulate organic carbon (POC)  
- microbial sources  
- marine algae  

These fields provide a first-order framework for interpreting organic matter provenance.

### Elliptical regions (Khan et al., 2019)

Elliptical regions are added to represent mangrove-specific endmembers:

- **Black mangroves (Avicennia germinans)**  
- **Red mangroves (Rhizophora mangle)**  

These regions are constructed with data from:

Khan et al. (2019), Marine Geology 415

and provide a more regionally constrained interpretation of mangrove-derived organic matter in Puerto Rico.

---

## Grouping and visualization

Samples are grouped using:`Identifier_5`

Each group is assigned:

- a unique color  
- a unique symbol  

Microbialite samples are explicitly highlighted to distinguish them from sediment core intervals.



## Outputs

### 1. Static publication plot

Object:`Sed_Org_Guanica`

Exported as:

- SVG (for Illustrator editing)  
- TIFF (high-resolution for publication)  

This figure is used in the *Sedimentology* manuscript.


### 2. Interactive plot

Object:`Interactive_Sed_Org_Guanica`


This version allows exploration of:

- depth information  
- sample grouping  
- geochemical values  

Exported as:`plots/Interactive_Sed_Org_Guanica.html`


---

## Reproducibility

Expected folder structure:
`StableIsotope_Plotting_Codes/`
`InputData/`
`Org_isotope_Guanica_data.csv`
`plots/`


Running the script will generate all outputs in the `plots/` directory.


## Notes

- All file paths are relative to the project directory  
- Missing values are removed prior to analysis to ensure data consistency  
- Color and symbol assignments are fixed for reproducibility  
- Environmental interpretation fields (rectangles and ellipses) are based on published frameworks and are used for visualization only  
- Final figure formatting may be refined in Adobe Illustrator; these edits do not modify data positions or values  


## References

Lamb, A. L., Wilson, G. P., & Leng, M. J. (2006). A review of coastal palaeoclimate and relative sea-level reconstructions using δ¹³C and C/N ratios in organic material. *Earth-Science Reviews*, 75, 29–57.

Khan, N. S., et al. (2019). The application of δ¹³C, TOC and C/N geochemistry of mangrove sediments to reconstruct Holocene paleoenvironments and relative sea levels, Puerto Rico. *Marine Geology*, 415.

# Script Set Up, Library calling and loading data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
setwd('/Users/bryanrodriguez/Library/Mobile Documents/com~apple~CloudDocs/PhD-PROJECT/R-codes/PRGeobio_XRD_Isotopes_Codes/StableIsotope_Plotting_Codes')

```



```{r packages, include = F, echo = FALSE}
library(ggplot2)      # Core plotting library for generating static figures
library(RColorBrewer) # Provides color palettes used for stratigraphic grouping
library(dplyr)        # Data manipulation and transformation (e.g., mutate, filtering)
library(plotly)       # Converts ggplot objects into interactive visualizations
library(htmlwidgets)  # Enables saving interactive plots as standalone HTML files
```

## Data import and preprocessing

Organic carbon and nitrogen data are imported from: `InputData/Org_isotope_Guanica_data.csv`


This dataset includes δ¹³Corg, TOC, TN, TOC/TN ratios, and associated metadata (e.g., depth and stratigraphic identifiers).

The column `DepthCore_m` is explicitly converted to numeric format to ensure compatibility with plotting and statistical analysis.

Rows containing missing values (e.g., low nitrogen concentrations or absent δ¹⁵N measurements) are removed to maintain consistency in geochemical interpretation.



```{r Load Organic Carbon and Nitrogen Data, echo = F}

#Organic Carbon and Nitrogen Data
Guanica_OrgData <- read.csv("InputData/Org_isotope_Guanica_data.csv")
IntergerColumns <- c("DepthCore_m")
Guanica_OrgData <- Guanica_OrgData %>% mutate_if(names(.) %in% IntergerColumns, as.numeric)

Guanica_OrgData <- na.omit(Guanica_OrgData) #excluding the ones that have low nitrogen concentration AND no d15N values
rm(IntergerColumns)

```


# Statistics

## Statistical analysis

This section evaluates whether the putative overwash layers (ZC3C) is geochemically distinct from other stratigraphic intervals.

A two-sample t-test is performed comparing:

- TOC/TN ratios  
- δ¹³Corg values  

between ZC3C samples and all remaining samples.

This analysis provides a quantitative assessment of differences in organic matter sources associated with the overwash events, if any. Thank you to Dr. Matthew Allen for the suggestion

```{r Statistical Significance test on Overswash event}
#subset the zone
ZC3C_Layer <- subset(Guanica_OrgData, Identifier_5 == "ZC3C")

#test the code
subset(ZC3C_Layer, Identifier_5 == "ZC3C")

#subset not in the zone
everyting_else <- subset(Guanica_OrgData, Identifier_5 != "ZC3C")

#test the code
subset(everyting_else, Identifier_5 == "ZC3C")

#Perform t-test
t.test(ZC3C_Layer$TOC_TN, everyting_else$TOC_TN)
t.test(ZC3C_Layer$d13C_org, everyting_else$d13C_org)

```


# Plotting

## Defined variables for plotting

### Mangrove reference fields (Khan et al., 2019)

Elliptical regions are defined to represent the isotopic and C/N ranges of mangrove-derived organic matter.

These fields correspond to:

- Black mangroves (*Avicennia germinans*)  
- Red mangroves (*Rhizophora mangle*)  

The ellipse geometry is constructed manually to approximate the ranges reported in Khan et al. (2019), providing a regional framework for interpreting mangrove contributions to sediment organic matter.

```{r Designing the ellipse regions}
ellipse_points <- function(cx, cy, rx, ry, n = 200){
  t <- seq(0, 2*pi, length.out = n)
  data.frame(
    x = cx + rx * cos(t),
    y = cy + ry * sin(t)
  )
}

ellipse_A <- ellipse_points(
  cx = 42,     # center x
  cy = -29,    # center y
  rx = 20,     # horizontal radius
  ry = 2
)
ellipse_A$type <- "BlackMangroves"

ellipse_B <- ellipse_points(
  cx = 60,
  cy = -26,
  rx = 18,
  ry = 1.5
)
ellipse_B$type <- "RedMangroves"

ellipse_df <- rbind(ellipse_A, ellipse_B) #Regions for the mangroves as shown in Khan et al., 2019 Red and black mangroves


```


### Plot styling and grouping scheme

Samples are grouped using the `Identifier_5` field, with each group assigned a unique color and symbol.

- C2 and C3 zones are represented using blue and orange palettes, respectively  
- Microbialite samples are highlighted separately  

Color and shape assignments are explicitly defined to ensure consistent visualization across plots and reproducibility between runs.

#### Environmental interpretation fields (Lamb et al., 2006)

Rectangular regions are defined to represent generalized organic matter sources based on δ¹³Corg and TOC/TN values.

These include:

- C₃ plants  
- C₄ plants  
- marine particulate organic carbon (POC)  
- microbial sources  
- marine algae  

These fields follow the framework described in Lamb et al. (2006) and provide a first-order interpretation of organic matter provenance.


```{r Defined variables for plotting}

# Generate distinct colors for C2 (blue shades) and C3 (orange shades)
set.seed(42)
c2_colors <- sample(brewer.pal(8, "Blues")[4:8], size = 5)
c3_colors <- sample(brewer.pal(8, "Oranges")[4:8], size = 4)

# Assign colors and shapes
zone_colors <- c(
  # C2 zones (shades of blue)
  "ZC2E" = c2_colors[1], "ZC2D" = "black", "ZC2C" = c2_colors[3], 
  "ZC2B" = "#4DAAE9", "ZC2A" = c2_colors[5],

  # C3 zones (shades of orange)
  "ZC3D" = "#8F3361", "ZC3C" = c3_colors[3], "ZC3B" = c3_colors[2], "ZC3A" = "#C65639",
  
  # Other groups (fixed colors)
  "Microbialite" = "green")

zone_shapes <- c(
  # C2 zones (unique shapes, new shape for ZC2A: diamond)
  "ZC2E" = 18, "ZC2D" = 20, "ZC2C" = 2, "ZC2B" = 1, "ZC2A" = 15,  # Diamond shape for ZC2A
  
  # C3 zones (unique shapes, keeping ZC3B the same)
  "ZC3D" = 8, "ZC3C" = 7, "ZC3B" = 6, "ZC3A" = 5,
  
  # Other groups (consistent shapes)
  "Microbialite" = 17)

# Rectangle data for regions
rect_data <- data.frame(
  xmin = c(4, 13, 5, 20, 6),
  xmax = c(7, 80, 10, 80, 9),
  ymin = c(-27, -30, -24, -17, -24.5),
  ymax = c(-12, -21, -18, 0, -16),
  zone = c("Microbial", "C3 plants", "Marine POC", "C4 plants", "Marine Algae")  # Custom labels
)


```

## Plotting

### Static isotope crossplot

This section generates the primary δ¹³Corg vs TOC/TN crossplot used in the manuscript.

Samples are plotted using predefined color and shape schemes, with environmental interpretation fields (rectangles and ellipses) overlaid for context.

The plot is formatted for publication, including axis limits, labeled units (VPDB), and minimal styling for clarity and post-processing in Illustrator.

```{r Plot for paper and for Illustrator}

#Plot d13C_org vs. TOC_TN with color-coded markers based on Identifier_1

Sed_Org_Guanica <- ggplot(Guanica_OrgData, aes(x = TOC_TN, y = d13C_org, color = Identifier_5, shape = Identifier_5, text =DepthCore_cm)) +
  geom_rect(
    data = rect_data,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = zone),
    inherit.aes = FALSE,
    alpha = 0.4
  ) +
  geom_point(size = 2.5, na.rm = TRUE) +
  scale_color_manual(values = zone_colors) +
  scale_shape_manual(values = zone_shapes) +
   scale_fill_manual(values = c(
     "Microbial" = "#A7C09E",
     "C3 plants" = "#D8AFA1",
     "Marine POC" = "#D9C8B0",
     "C4 plants" = "#BAAED1",
     "Marine Algae" = "lightblue",
     "BlackMangroves" = "black",
     "RedMangroves" = "red"
   )) +
  
 geom_polygon(
  data = ellipse_df,
  aes(x = x, y = y, group = type, fill = type),
  alpha = 0.15,
  color = NA,
  inherit.aes = FALSE
) +
  
  geom_point(size = 3.5, na.rm = T) +
  scale_color_manual(values = zone_colors) +
  scale_shape_manual(values = zone_shapes) +
  xlab(expression(paste(C[phantom(0)][org],"/",N[phantom(0)][total], " ", "(‰, VPDB)"))) +  # X-axis label
  ylab(expression(paste(delta^{13}, C[phantom(0)][org], " ", "(‰, VPDB)"))) +  # Y-axis label
  ggtitle(expression(paste(delta^{13}, C[phantom(0)][org], " ", "and", " ", C[phantom(0)][org],"/",N[phantom(0)][total], " ", "Values of sediment cores and microbialites in Laguna Providencia and Montalva, Guánica, PR" ))) +
  theme_minimal() + # Uses a minimal theme for aesthetics

theme(legend.position = "right", legend.key.size = unit(0.5, "cm"), legend.title = element_text(hjust = 0.5, size = 12),  legend.text = element_text(hjust = 0.5), plot.title = element_text(hjust = 0.3, size = 9, face = "bold"), plot.background =   element_rect(fill = "white"), axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14), panel.grid = element_blank(), panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5)) +
  
scale_x_continuous(breaks = seq(0, 70, by = 5)) + # Labels at every 1 interval
scale_y_continuous(breaks = seq(-28, -10, by = 2)) + # Labels at every 1 interval
coord_cartesian(xlim = c(0, 70), ylim = c(-28, -10)) +
labs(color = "Legend") +
labs(shape = "Legend") +
labs(fill = "Legend") +
#guides(fill=guide_legend(title="Regions"))
guides(fill="none")
Sed_Org_Guanica
print("Plot ready to be modified on Illustrator")

```

## Interactive plot

An interactive version of the δ¹³Corg vs TOC/TN plot is generated using `plotly`.

This version allows exploration of individual data points, including:

- depth information  
- sample grouping  
- geochemical values  

It is intended for exploratory analysis and is exported as a standalone HTML file.

```{r Make interactive plot}

Sed_Org_Guanica <-  ggplot(Guanica_OrgData, aes(x = TOC_TN, y = d13C_org, color = Identifier_5, shape = Identifier_5, text =DepthCore_cm)) +
  geom_rect(
    data = rect_data,
    aes(xmin = xmin, xmax = xmax, ymin = ymin, ymax = ymax, fill = zone),
    inherit.aes = FALSE,
    alpha = 0.4
  ) +
  geom_point(size = 2.5, na.rm = T) +
  scale_color_manual(values = zone_colors) +
  scale_shape_manual(values = zone_shapes) +
   scale_fill_manual(values = c(
     "Microbial" = "#A7C09E",
     "C3 plants" = "#D8AFA1",
     "Marine POC" = "#D9C8B0",
     "C4 plants" = "#BAAED1",
     "Marine Algae" = "lightblue",
     "BlackMangroves" = "black",
     "RedMangroves" = "red"
   )) +
  
   geom_polygon(
  data = ellipse_df,
  aes(x = x, y = y, group = type, fill = type),
  alpha = 0.15,
  color = NA,
  inherit.aes = FALSE
) +
  
  geom_point(size = 2.5, na.rm = T) +
  scale_color_manual(values = zone_colors) +
  scale_shape_manual(values = zone_shapes) +
  ylab("d13Corg") +  # Set y-axis label
  xlab("TOC/TN") +
  ggtitle("Corg and TOC/TN Values of sediment cores and microbialites in Laguna Providencia and Montalva, Guánica, PR" ) +
  theme_minimal() +
   theme(legend.position = "right", legend.key.size = unit(0.5, "cm"), legend.title = element_text(hjust = 0.5, size = 12),  legend.text = element_text(hjust = 0.5), plot.title = element_text(hjust = 0, size = 10), axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14), plot.background =   element_rect(fill = "white"), panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5)) +
  scale_x_continuous(breaks = seq(0, 70, by = 5)) + # Labels at every 1 interval
scale_y_continuous(breaks = seq(-28, -10, by = 2)) + # Labels at every 1 interval
coord_cartesian(xlim = c(0, 70), ylim = c(-28, -10)) +
  labs(color = "Legend") +
labs(shape = "Legend") +
labs(fill = "Legend")


#Make ggplotly
Interactive_Sed_Org_Guanica <- ggplotly(Sed_Org_Guanica, tooltip = c("x", "y", "color", "text"))

Interactive_Sed_Org_Guanica



```

## Output export

Figures are exported in multiple formats:

- **HTML**: interactive visualization  
- **SVG**: vector format for editing in Adobe Illustrator  
- **TIFF**: high-resolution format for journal submission  

All outputs are saved in the `plots/` directory.

Final formatting adjustments may be performed in Illustrator without altering data positions or underlying values.

```{r Save plots}

# Save as an interactive HTML file
saveWidget(Interactive_Sed_Org_Guanica, "plots/Interactive_Sed_Org_Guanica.html", selfcontained = T, libdir = NULL)
print("Interactive HTML file Interactive_Sed_Org_Guanica saved")


#Save as SVG (for Illustrator editing)
ggsave("plots/Sed_Org_Guanica.svg", Sed_Org_Guanica, width = 6, height = 6, units = "in", device = "svg")
print("SVG file saved as Sed_Org_Guanica.svg")


# Save as TIFF
ggsave(filename = "plots/Sed_Org_Guanica_KhanAddition.tiff", plot = Sed_Org_Guanica, device = "tiff", width = 13, height = 7, dpi = 600, compression = "lzw")
print("TIFF file saved as Sed_Org_Guanica.tiff")

```






