---
title: 'Code 1: isoCarbs'
author: "Bryan Rodriguez-Colon"
date: "2025-03-23"
output: html_document
---

***Disclaimer: This code was designed by Bryan Rodriguez-Colon with the help of chatGPT-5.***

# Overview

This script generates carbonate stable isotope crossplots (δ¹³C vs δ¹⁸O) for sediment and carbonate samples collected from the Guánica coastal lagoon system, southwestern Puerto Rico.

The code reads carbonate isotope data, defines color and symbol schemes corresponding to stratigraphic zones and sample groups, and produces two outputs:

1. A publication-quality static plot used in the manuscript  
2. An interactive plot for exploratory data visualization  

The final figures are exported in formats suitable for publication and figure editing.

## Associated manuscript

Rodríguez-Colón, B. J., et al. (2026)  
*Paleoenvironmental and Regional Controls on Microbialite Development and Distribution in a Coastal Lagoon of Southwestern Puerto Rico*  
Sedimentology

**Note:** The plots generated in this script are not included in the associated manuscript but are presented in the dissertation appendix corresponding to this chapter, where they support additional interpretation of carbonate isotope variability. The microbialite, beach sand, and Ponce Limestone datasets included in these plots are published in Rodríguez-Colón et al. (in review) in *Scientific Reports*.

## Input data

The script reads carbonate isotope data from: `Isotope_Data_Guanica.csv`


The dataset contains measurements of:

- δ¹³C (‰ VPDB)  
- δ¹⁸O (‰ VPDB)  

as well as metadata fields used for plotting:

| Column         | Description                          |
|----------------|--------------------------------------|
| Identifier_2   | Sample identifier                    |
| Identifier_3   | Sample group or stratigraphic zone   |
| Mineralogy     | Dominant mineral phase               |
| DepthCore_m    | Sample depth (meters)                |



## Plot structure

The plots generated in this script show:

**X-axis**  
δ¹⁸O (‰ VPDB)

**Y-axis**  
δ¹³C (‰ VPDB)

These isotope values are widely used in carbonate sedimentology to interpret:

- evaporative conditions  
- diagenetic alteration  
- mixing between marine and restricted waters  


## Color and symbol scheme

Samples are grouped using: `Identifier_3`


Each group is assigned:

- a unique color  
- a unique symbol  

### Stratigraphic groupings

**C2 zones**  
Displayed using blue color gradients.

**C3 zones**  
Displayed using orange color gradients.

### Additional sample groups

- Ponce Limestone mixed samples  
- sand samples  
- microbialite samples  

These groupings correspond to stratigraphic and sedimentological units discussed in the manuscript.


## Outputs

### 1. Static publication plot

Object name: `Sed_Carbs_Guanica`


Exported as:

**SVG (vector format for Illustrator)**  
`plots/Sed_Carbs_Guanica.svg`


**TIFF (high-resolution for journal submission)**  
`plots/Sed_Carbs_Guanica.tiff`


### 2. Interactive plot

Object name: `Interactive_Sed_Carbs_Guanica`


Includes hover information:

- sample identifier  
- mineralogy  
- isotope values  

Exported as:`plots/Interactive_Sed_Carbs_Guanica.html`




## Reproducibility

Expected folder structure:

`StableIsotope_Plotting_Codes/`
`InputData/Isotope_Data_Guanica.csv`
`plots/`


Running the script will generate all figures in the `plots/` directory.


## Notes

- All file paths are relative to the project directory  
- The script assumes cleaned isotope data with consistent formatting  
- Color and symbol assignments are manually defined to match manuscript figures  
- Final figure formatting (e.g., label placement, font adjustments, panel alignment) may be refined in Adobe Illustrator; these edits do not alter data positions or the underlying dataset  


# Script Set Up, Library calling and loading data

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
#setwd('/Users/bryanrodriguez/Library/Mobile Documents/com~apple~CloudDocs/PhD-PROJECT/R-codes/PRGeobio_XR_C14_StableIsotopes_Codes/StableIsotope_Plotting_Codes')

```



```{r packages, include = T, message = F}
library(ggplot2)      # Core plotting library for generating static figures
library(RColorBrewer) # Provides color palettes used for stratigraphic grouping
library(dplyr)        # Data manipulation and transformation (e.g., mutate, filtering)
library(plotly)       # Converts ggplot objects into interactive visualizations
library(htmlwidgets)  # Enables saving interactive plots as standalone HTML files
```


## Data import and preprocessing

Carbonate stable isotope data are imported from: `InputData/Isotope_Data_Guanica.csv`


This dataset contains δ¹³C and δ¹⁸O measurements along with associated metadata (e.g., sample identifiers, mineralogy, and depth).

### Data type correction

The column:`DepthCore_m` is explicitly converted to numeric format to ensure compatibility with downstream analyses and plotting functions.

This step addresses potential issues arising from inconsistent data types (e.g., character values resulting from CSV import), which can interfere with quantitative operations and axis scaling.


```{r Load data}

#Carbonate Data
Guanica_CarbData <- read.csv("InputData/Isotope_Data_Guanica.csv")

#Fix interger chr in depth column
IntergerColumns <- c("DepthCore_m")
Guanica_CarbData <- Guanica_CarbData %>% mutate_if(names(.) %in% IntergerColumns, as.numeric)


```

# Plotting

## Plot styling and grouping scheme

This section defines the color and symbol mappings used to visualize carbonate stable isotope data according to stratigraphic zones and sample group classifications.

### Color assignment

Distinct color palettes are generated to differentiate major stratigraphic groupings:

- **C2 zones** are represented using shades of blue  
- **C3 zones** are represented using shades of orange  

A fixed random seed (`set.seed(42)`) is used to ensure reproducibility of the sampled color subsets from the RColorBrewer palettes. Selected colors are drawn from mid-to-dark ranges of each palette to enhance visual contrast and readability.

Additional sample groups are assigned fixed colors to maintain consistency across figures:

- Ponce Limestone samples (mixed and treated)  
- sand samples  
- microbialite samples  

These assignments ensure consistent visual representation across all plots in the project.

### Shape assignment

Each sample group (defined by `Identifier_3`) is also assigned a unique plotting symbol to allow simultaneous differentiation by both color and shape.

- **C2 and C3 zones** are assigned distinct shapes to distinguish stratigraphic units  
- **Ponce Limestone samples** are represented by filled circles  
- **Sand samples** use a distinct symbol to separate them from carbonate samples  
- **Microbialite samples** are represented by filled triangles  

This dual encoding (color + shape) improves interpretability, particularly in grayscale reproduction or for color-vision deficiencies.

```{r Defined variables for plotting}

# Generate distinct colors for C2 (blue shades) and C3 (orange shades)
set.seed(42)
c2_colors <- sample(brewer.pal(8, "Blues")[4:8], size = 5)
c3_colors <- sample(brewer.pal(8, "Oranges")[4:8], size = 4)

# Assign colors and shapes
zone_colors <- c(
  # C2 zones (shades of blue)
  "ZC2E" = c2_colors[1], "ZC2D" = "black", "ZC2C" = c2_colors[3], 
  "ZC2B" = "#4DAAE9", "ZC2A" = c2_colors[5],

  # C3 zones (shades of orange)
  "ZC3D" = "#8F3361", "ZC3C" = c3_colors[3], "ZC3B" = c3_colors[2], "ZC3A" = "#C65639",
  
  # Other groups (fixed colors)
  "PLDolMixed" = "purple", "PLCalMixed" = "red", "PLDolTreated" = "orchid",
  "SandMixed" = "grey28",
  "PS_Top" = "green", "PS_BlackL" = "darkgreen"
)

zone_shapes <- c(
  # C2 zones (unique shapes, new shape for ZC2A: diamond)
  "ZC2E" = 18, "ZC2D" = 20, "ZC2C" = 2, "ZC2B" = 1, "ZC2A" = 15,  # Diamond shape for ZC2A
  
  # C3 zones (unique shapes, keeping ZC3B the same)
  "ZC3D" = 8, "ZC3C" = 7, "ZC3B" = 6, "ZC3A" = 5,
  
  # Other groups (consistent shapes)
  "PLDolMixed" = 19, "PLCalMixed" = 19, "PLDolTreated" = 19, # Circles for PL
  "SandMixed" = 12,
  "PS_Top" = 17, "PS_BlackL" = 17 # Filled triangles for microbialites
)



```

## Generating Plot

### Static isotope crossplot

This section generates the primary carbonate isotope crossplot (δ¹³C vs δ¹⁸O) using `ggplot2`, with samples grouped by `Identifier_3` and styled using predefined color and shape schemes.

The plot is formatted for publication, including labeled axes (VPDB), defined axis limits, and reference lines at δ¹³C = 0 and δ¹⁸O = 0. Visual styling is optimized for clarity and subsequent editing in Illustrator.

An additional block is included below to generate an interactive version of this plot for exploratory analysis.

```{r Oficial plot by Zones for paper}


# Plot
Sed_Carbs_Guanica <- ggplot(Guanica_CarbData, aes(x = d18O, y = d13C, color = Identifier_3, shape = Identifier_3)) +
  geom_point(size = 3.5, stroke = 1, na.rm = TRUE) + # Black stroke for clarity
  scale_color_manual(values = zone_colors) +
  scale_shape_manual(values = zone_shapes) +
  xlab(expression(paste(delta^{18}, O[phantom(0)][carb], " ", "(‰, VPDB)"))) +  # X-axis label
  ylab(expression(paste(delta^{13}, C[phantom(0)][carb], " ", "(‰, VPDB)"))) +  # Y-axis label
  ggtitle(expression(paste(delta^{13}, C[phantom(0)][carb], " ", "and", " ", delta^{18}, O[phantom(0)][carb], " ", "Values of different samples in Laguna Providencia and Montalva, Guánica, PR" ))) +
  theme_minimal() +
  theme(
    legend.position = "right", 
    legend.key.size = unit(0.5, "cm"), 
    legend.title = element_text(hjust = 0.5, size = 12),  
    legend.text = element_text(hjust = 0.5), 
    plot.title = element_text(hjust = 0.15, size = 10, face = "bold"), 
    axis.title.x = element_text(size = 14),
    axis.title.y = element_text(size = 14),
    plot.background = element_rect(fill = "white"),
    panel.grid = element_blank(),         # Remove all grid lines
    panel.border = element_rect(          # Add a square border
      colour = "black", fill = NA, linewidth = 0.5
    )
  ) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "gray40") + # Solid axis lines
  geom_vline(xintercept = 0, linetype = "dashed", color = "gray40") +
  scale_x_continuous(breaks = seq(-6, 3, by = 1)) + 
  scale_y_continuous(breaks = seq(-12, 5, by = 1)) +
  coord_cartesian(xlim = c(-6, 3), ylim = c(-12, 5)) + # Adjust plot limits
  labs(color = "Legend") +
  labs(shape = "Legend") +
  labs(fill = "Legend")





# Display the plot
Sed_Carbs_Guanica

print("Plot ready to be modified on Illustrator")



```

### Interactive Plot

An interactive version is generated using `plotly`.

This version allows exploration of individual data points, including:

- depth information  
- sample grouping  
- geochemical values
- Main minerology

It is intended for exploratory analysis and is exported as a standalone HTML file.

```{r Make interactive plot}

Interactive_Sed_Carbs_Guanica <- ggplot(Guanica_CarbData, aes(x = d18O, y = d13C, color = Identifier_3, shape = Identifier_3, text =Identifier_2, label = Mineralogy)) +
  geom_point(size = 2.5, na.rm = T) +
  scale_color_manual(values = zone_colors) +
  scale_shape_manual(values = zone_shapes) +
  xlab("d18O carb") +  # Set x-axis label
  ylab("d13C carb") +  # Set y-axis label
    ggtitle("Carbonate Isotope values across Laguna Providencia, Guánica, PR" ) +
  theme_minimal() +
  theme(legend.position = "right", legend.key.size = unit(0.5, "cm"), legend.title = element_text(hjust = 0.5, size = 12),  legend.text = element_text(hjust = 0.5), plot.title = element_text(hjust = 0.15, size = 10), axis.title.x = element_text(size = 14), axis.title.y = element_text(size = 14), plot.background =   element_rect(fill = "white"), panel.border = element_rect(colour = "black", fill = NA, linewidth = 0.5)) +
  geom_hline(yintercept = 0, linetype = "solid", color = "gray40") + # Solid axis lines
  geom_vline(xintercept = 0, linetype = "solid", color = "gray40") + # Solid axis lines
  scale_x_continuous(breaks = seq(-6, 3, by = 1)) + # Labels at every 1 interval
  scale_y_continuous(breaks = seq(-12, 5, by = 1)) + # Labels at every 1 interval
  coord_cartesian(xlim = c(-6, 3), ylim = c(-12, 5)) +
  labs(color = "Legend") +
  labs(shape = "Legend") +
  labs(fill = "Legend") +
  guides(fill=guide_legend(title="Legend"))

#Make ggplotly
Interactive_Sed_Carbs_Guanica <- ggplotly(Interactive_Sed_Carbs_Guanica, tooltip = c("x", "y", "color", "text", "label"))

Interactive_Sed_Carbs_Guanica

print("Interactive plot ready for inspection")

```

## Output export

Figures are exported in multiple formats:

- **HTML**: interactive visualization  
- **SVG**: vector format for editing in Adobe Illustrator  
- **TIFF**: high-resolution format for journal submission  

All outputs are saved in the `plots/` directory.

Final formatting adjustments may be performed in Illustrator without altering data positions or underlying values.

```{r Saving Plots}


# Save as an interactive HTML file
saveWidget(Interactive_Sed_Carbs_Guanica, "plots/Interactive_Sed_Carbs_Guanica.html", selfcontained = T, libdir = NULL)
print("Interactive HTML file Interactive_Sed_Carbs_Guanica saved")


#Save as SVG (for Illustrator editing)
ggsave("plots/Sed_Carbs_Guanica.svg", Sed_Carbs_Guanica, width = 6, height = 6, units = "in", device = "svg")
print("SVG file saved as Sed_Carbs_Guanica.svg")


# Save as TIFF
ggsave(filename = "plots/Sed_Carbs_Guanica.tiff", plot = Sed_Carbs_Guanica, device = "tiff", width = 13, height = 13, dpi = 600, compression = "lzw")
print("TIFF file saved as Sed_Carbs_Guanica.tiff")

```




